---
title: "Phyloseq_artifact_generation and Quality_Control 02122026"
output: html_notebook
---

## Formatting output from DADA2 in Qiime2 pipeline for Phyloseq 

```{r}
library(reshape)
library(plyr)
library(dplyr)
library(dplyr)
library(tidyr)
library(tidyverse)
library(reshape2)
library(ggpubr)
library(phyloseq)
library(here)
```


```{r}
# -----------------------------
# 1. Read taxonomy table
# -----------------------------

tax <- read.delim(
  file.path(base_dir, "taxonomy", "taxonomy.tsv"),
  sep = "\t",
  header = TRUE,
  stringsAsFactors = FALSE
)

tax_split <- tax %>%
  separate(
    Taxon,
    into = c("Kingdom","Phylum","Class","Order","Family","Genus","Species"),
    sep = ";",
    fill = "right"
  ) %>%
  mutate(across(Kingdom:Species, ~gsub("^.*__", "", .))) %>%
  column_to_rownames("Feature.ID") %>%
  select(Kingdom:Species)


tax_table_ps <- tax_table(as.matrix(tax_split))
```


```{r}
# -----------------------------
# 2. Read OTU / ASV table
# -----------------------------

library(ape)
library(biomformat)

# --- 2️⃣ Read BIOM file manually ---
biom_path <- "/Users/maxmiao/Library/CloudStorage/Box-Box/Lankau_Lab/Ch4_Greenhouse_Exmpt_2018/qiime2_16S/16S_02122026/table/feature-table.biom"
biom_file <- read_biom(biom_path)

# Convert to matrix
otu_mat <- as(biom_data(biom_file), "matrix")

# Ensure numeric (QIIME2 sometimes exports float or characters)
mode(otu_mat) <- "numeric"

# Convert to phyloseq OTU table
otu_table_ps <- otu_table(otu_mat, taxa_are_rows = TRUE)

# Quick check
otu_table_ps[1:5, 1:5]

```


```{r}
# -----------------------------
# 3. Read sample metadata
# -----------------------------
meta_df <- read_csv("metadata3.csv")


meta <- meta_df %>%
  column_to_rownames("Submission_ID") %>%
  sample_data()
```


```{r}
# -----------------------------
# 4. Read phylogeny tree
# -----------------------------
tree <- read_tree(
  file.path(base_dir, "tree", "tree.nwk")
)

phy_tree_ps <- phy_tree(tree)
is.rooted(phy_tree_ps)
```


##Generate as Phyloseq object: 

```{r}
ps <- merge_phyloseq(
  otu_table_ps,
  tax_table_ps,
  meta,
  phy_tree_ps
)

ps

##Generate rarefraction curves for pruning: 
```

# 1. Extract the OTU/ASV table as a standard matrix
otu_mat <- as(otu_table(ps_clean), "matrix")

# 2. Vegan requires samples to be ROWS. If they are columns, transpose the matrix:
if (taxa_are_rows(ps_clean)) {
  otu_mat <- t(otu_mat)
}

# 3. Plot the rarefaction curves
rarecurve(otu_mat, 
          step = 250,             # Step size for calculation (lower = smoother but slower)
          col = "blue",           # Color of lines
          label = FALSE,          # Hide sample name labels to avoid visual clutter
          xlab = "Read Depth", 
          ylab = "Observed Taxa")

# 4. Add a vertical line at a candidate cutoff (e.g., 5000 reads)
abline(v = 5000, col = "red", lty = 2, lwd = 2)
```


##Quality Control(s)

```{r}
ps_clean <- ps %>%
  subset_taxa(
    !Family %in% c("Mitochondria", "Chloroplast") &
    !Order  %in% c("Mitochondria", "Chloroplast")
  )

# Convert Family and Order to lower case for consistent filtering
ps_clean <- subset_taxa(
  ps_clean,
  !tolower(Family) %in% c("mitochondria", "chloroplast") &
  !tolower(Order)  %in% c("mitochondria", "chloroplast")
)
ps1_prune <- prune_samples(reads >= 5000, ps_clean)

#Export for alpha diversity analysis: 
#saveRDS(ps_clean,"ps1_AlphaDiv_02182026.RDS") #Use for Alpha Diversity

```

## Further Quality Control for DESeq2 and Beta Diversity:
```{r}
# ===========================
# 1. Remove singletons (ASVs in only 1 sample)
# ===========================
ps_filtered <- filter_taxa(ps1_prune, function(x) sum(x > 0) > 1, TRUE)

# ===========================
# 2. Apply prevalence filter
#    Keep ASVs present in at least 10% of samples 
# ===========================
prevalenceThreshold <- 0.1
sampleCountThreshold <- ceiling(prevalenceThreshold * nsamples(ps_filtered))

ps_filtered <- filter_taxa(
  ps_filtered,
  function(x) sum(x > 0) >= sampleCountThreshold,
  TRUE
)

ps_desq = ps_filtered # singleton + prevalence filtering only

#saveRDS(ps_desq,"ps1_DESeq2_02232026.RDS") #Use for DESEQ2 analysis

# ===========================
# 3. Apply relative abundance filter
#    Keep ASVs that make up >0.25% of total reads in any sample. For ASV filtering based on "0.25% in any sample", you need sample-wise relative abundance.
# ===========================
ps.rel <- transform_sample_counts(
  ps_filtered,
  function(x) x / sum(x)
)

ps_filtered <- filter_taxa(
  ps.rel,
  function(x) any(x > 0.0025),
  TRUE
)

keep_taxa <- taxa_names(
  filter_taxa(
    ps.rel,
    function(x) any(x > 0.0025),
    TRUE
  )
)

ps_filtered <- prune_taxa(keep_taxa, ps_filtered)
#This preserves the original counts while filtering based on relative abundance.

ps.beta = ps_filtered   # singleton + prevalence + abundance filtering
#saveRDS(ps.beta,"ps1_BetaDiv_02182026.RDS")

# ===========================
# Note: Using 'any(x / sum(x) > abundanceThreshold)' what we initially did calculates relative abundance within each taxon across all samples, not within each sample. We made adjustment of ASV filtering based on at least 0.25% in any sample. 
```

