---
title: "Phyloseq_artifact_generation and Quality_Control 02122026"
output: html_notebook
---

## Formatting output from DADA2 in Qiime2 pipeline for Phyloseq 

```{r}
library(reshape)
library(plyr)
library(dplyr)
library(dplyr)
library(tidyr)
library(tidyverse)
library(reshape2)
library(ggpubr)
library(phyloseq)
library(here)
```


```{r}
# -----------------------------
# 1. Read taxonomy table
# -----------------------------

tax <- read.delim(
  file.path(base_dir, "taxonomy", "taxonomy.tsv"),
  sep = "\t",
  header = TRUE,
  stringsAsFactors = FALSE
)

tax_split <- tax %>%
  separate(
    Taxon,
    into = c("Kingdom","Phylum","Class","Order","Family","Genus","Species"),
    sep = ";",
    fill = "right"
  ) %>%
  mutate(across(Kingdom:Species, ~gsub("^.*__", "", .))) %>%
  column_to_rownames("Feature.ID") %>%
  select(Kingdom:Species)


tax_table_ps <- tax_table(as.matrix(tax_split))
```


```{r}
# -----------------------------
# 2. Read OTU / ASV table
# -----------------------------

library(ape)
library(biomformat)

# --- 2️⃣ Read BIOM file manually ---
biom_path <- "/Users/maxmiao/Library/CloudStorage/Box-Box/Lankau_Lab/Ch4_Greenhouse_Exmpt_2018/qiime2_16S/16S_02122026/table/feature-table.biom"
biom_file <- read_biom(biom_path)

# Convert to matrix
otu_mat <- as(biom_data(biom_file), "matrix")

# Ensure numeric (QIIME2 sometimes exports float or characters)
mode(otu_mat) <- "numeric"

# Convert to phyloseq OTU table
otu_table_ps <- otu_table(otu_mat, taxa_are_rows = TRUE)

# Quick check
otu_table_ps[1:5, 1:5]

```


```{r}
# -----------------------------
# 3. Read sample metadata
# -----------------------------
meta_df <- read_csv("metadata3.csv")


meta <- meta_df %>%
  column_to_rownames("Submission_ID") %>%
  sample_data()
```


```{r}
# -----------------------------
# 4. Read phylogeny tree
# -----------------------------
tree <- read_tree(
  file.path(base_dir, "tree", "tree.nwk")
)

phy_tree_ps <- phy_tree(tree)
is.rooted(phy_tree_ps)
```


##Generate as Phyloseq object: 

```{r}
ps <- merge_phyloseq(
  otu_table_ps,
  tax_table_ps,
  meta,
  phy_tree_ps
)

ps

```


##Quality Control(s)

```{r}
ps_clean <- ps %>%
  subset_taxa(
    !Family %in% c("Mitochondria", "Chloroplast") &
    !Order  %in% c("Mitochondria", "Chloroplast")
  )

# Convert Family and Order to lower case for consistent filtering
ps_clean <- subset_taxa(
  ps_clean,
  !tolower(Family) %in% c("mitochondria", "chloroplast") &
  !tolower(Order)  %in% c("mitochondria", "chloroplast")
)
ps1_prune <- prune_samples(reads >= 8000, ps_clean)

#Export for alpha diversity analysis: 
#saveRDS(ps_clean,"ps1_AlphaDiv_02182026.RDS") #Use for Alpha Diversity

```


```{r}
# Apple 0.25% threshold (for beta diversity)
ps_relabund <- transform_sample_counts(ps_clean, function(x) x / sum(x))
keep_taxa <- taxa_names(ps_relabund)[apply(otu_table(ps_relabund), 1, function(x) any(x > 0.0025))]
ps_filtered <- prune_taxa(keep_taxa, ps_clean)

# Drop Zero abundance taxa (for beta diversity)
ps_filtered <- prune_taxa(taxa_sums(ps_filtered) > 0, ps_filtered)
```


```{r}
#saveRDS(ps_filtered,"ps1_filtered_02182026.RDS") #Use for Beta_diversity
```


#Quality Control for DESeq2:
```{r}
# ===========================
# 1. Remove singletons (ASVs in only 1 sample)
# ===========================
physeq_filtered <- filter_taxa(ps_clean, function(x) sum(x > 0) > 1, TRUE)

# ===========================
# 2. Apply prevalence filter
#    Keep ASVs present in at least 10% of samples
# ===========================
prevalenceThreshold <- 0.1
sampleCountThreshold <- prevalenceThreshold * nsamples(physeq_filtered)
physeq_filtered <- filter_taxa(physeq_filtered, function(x) sum(x > 0) >= sampleCountThreshold, TRUE)

# ===========================
# 3. Apply relative abundance filter
#    Keep ASVs that make up >0.25% of total reads in any sample
# ===========================
abundanceThreshold <- 0.0025
physeq_filtered <- filter_taxa(physeq_filtered, function(x) any(x / sum(x) > abundanceThreshold), TRUE)

#saveRDS(physeq_filtered,"ps1_DESeq2_02232026.RDS") #Use for DESEQ2 analysis

```



Check read counts: 
```{r}
library(data.table)
ps1 = readRDS("ps1_filtered_02182026.RDS")

readcount = data.table(as(sample_data(ps1), "data.frame"),
                 TotalReads = sample_sums(ps1), 
                 keep.rownames = TRUE)
setnames(readcount, "rn", "Submission_ID")
ggplot(readcount, aes(TotalReads)) + geom_histogram() + ggtitle("Sequencing Depth")
```
As we can see we have lots of samples that have not many reads...we need to find a good cutoff. We will now generate a rarefraction plot 




```{r}
#First need to remove samples with zero reads:
ps1.clean <- prune_samples(sample_sums(ps1) > 0, ps1)

readcount = data.table(as(sample_data(ps1.clean), "data.frame"),
                 TotalReads = sample_sums(ps1.clean), 
                 keep.rownames = TRUE)
setnames(readcount, "rn", "Submission_ID")
ggplot(readcount, aes(TotalReads)) + geom_histogram() + ggtitle("Sequencing Depth")
```


So it turns out we need to do more
```{r}
library(phyloseq)

rarecurve(
  t(otu_table(ps1.clean)),
  step = 100,
  sample = min(sample_sums(ps1.clean)),
  col = "grey",
  label = FALSE
)


```

#Diagnosing issue related to rarefraction: 
```{r}
otu_mat <- as(otu_table(ps1), "matrix")
if (taxa_are_rows(ps1)) otu_mat <- t(otu_mat)
```


```{r}
rowSums(otu_mat == 0)
rowSums(otu_mat > 0)
rowSums(otu_mat)
```


```{r}
bad_samples <- rownames(otu_mat)[rowSums(otu_mat > 0) < 2]
bad_samples
```


```{r}
ps1_clean <- prune_samples(
  sample_names(ps1) %in% setdiff(sample_names(ps1), bad_samples),
  ps1
)
```


```{r}
otu_mat <- as(otu_table(ps1_clean), "matrix")
if (taxa_are_rows(ps1_clean)) otu_mat <- t(otu_mat)

rarecurve(
  otu_mat,
  step   = 100,
  sample = min(rowSums(otu_mat)),
  col    = "grey",
  label  = FALSE
)
```
So based on this; our cutoff will be 8000 for downstream analysis 

```{r}
ps1_filt <- prune_samples(reads >= 8000, ps1)
```
Notice that it originally had 356 we have 224 left.  

```{r}
#saveRDS(ps1_filt,"ps1_Clean_02182026.RDS")
```

