---
title: "DESeq2_yield_unfiltered"
output: html_notebook
---

Goal of Differential Analysis: 

1) Examine differential abundance of microbial communities among Positive vs Negative selected lines. 

2) Examine differential abundance of microbial communities among positive vs negative plant responsive samples. 

3) What taxa are overlapping between (1) and (2)?


```{r}
library(phyloseq)
library(tidyverse)
library(vegan)
library(ggplot2)
library(lme4)
library(lmerTest)
library(emmeans)
library(DESeq2)
```

Quality Control steps performed in DESeq2_Clade.Rmd and saved as "ps1.RDS"
```{r}
physeq_trimmed = readRDS(file = "ps_merged.RDS")
ps1 = readRDS("ps1.RDS")
#physeq_glom <- tax_glom(physeq_trimmed, taxrank = "Genus", NArm = F)
# Filter out taxa found in less than 10% of samples

#physeq_trimmed = physeq_filter_taxa
#physeq_trimmed
```
After filtering: we removed taxa not seen more than 1 times in at least 10% of the samples. This protects against an OTU with small mean & trivially large C.V. We only end up with 532 Features... 


**Functions to be used:**

```{r}
#Function 1:
geomMeans_DESeq2_Res = function(ps_object){
diagdds_clade= phyloseq_to_deseq2(ps_object, ~ Response)

gm_mean = function(x, na.rm=TRUE){
  exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
}

geoMeans = apply(counts(diagdds_clade), 1, gm_mean)
diagdds = estimateSizeFactors(diagdds_clade, geoMeans = geoMeans)

diagdds_clade = DESeq(diagdds, fitType="local")
return(diagdds_clade)
}

geomMeans_DESeq2_ResT = function(ps_object){
diagdds_clade= phyloseq_to_deseq2(ps_object, ~ Response*Iteration)

gm_mean = function(x, na.rm=TRUE){
  exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
}

geoMeans = apply(counts(diagdds_clade), 1, gm_mean)
diagdds = estimateSizeFactors(diagdds_clade, geoMeans = geoMeans)

diagdds_clade = DESeq(diagdds, fitType="local")
return(diagdds_clade)
}



#Function 2:
# Calculate size factors using modified geometric means that can handle zeros and estimate normalization factors

geomMeans_DESeq2_Select = function(ps_object){
diagdds_clade= phyloseq_to_deseq2(ps_object, ~ Selection.x * Iteration)

gm_mean = function(x, na.rm=TRUE){
  exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
}

geoMeans = apply(counts(diagdds_clade), 1, gm_mean)
diagdds = estimateSizeFactors(diagdds_clade, geoMeans = geoMeans)

diagdds_clade = DESeq(diagdds, fitType="local")
return(diagdds_clade)
}

#Function 3: significant taxa < alpha
DESeq_table = function(DESeq2_ds,ps_object){
res = results(DESeq2_ds, cooksCutoff = FALSE)
alpha = 0.05
sigtab = res[which(res$padj < alpha), ]
sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(ps_object)[rownames(sigtab), ], "matrix"))
return(sigtab)
}

#Function 4: All results

res_table = function(diagdds_clade,ps_object){
  res = results(diagdds_clade, cooksCutoff = FALSE)
  res_A_Phyla = cbind(as(res, "data.frame"), as(tax_table(ps_object)[rownames(res), ], "matrix"))
}

```


##Aim 1: Quantitative relative analysis of community structure to response outcome 


###At Genus level:
```{r}
#Andigenum and Response:
physeq_Andigenum = subset_samples(ps1, Clade == 'A' & Response != "NA" & Soil_inoculation != "Sterile")
physeq_Andigenum <- tax_glom(physeq_Andigenum, taxrank = "Genus", NArm = F)
physeq_Andigenum = subset_samples(physeq_Andigenum, taxa_sums(physeq_Andigenum) > 0)
physeq_Andigenum = subset_samples(physeq_Andigenum, sample_sums(physeq_Andigenum) > 0)
diagdds_ResA_ASV = geomMeans_DESeq2_Res(physeq_Andigenum)
res_A_ASV= res_table(diagdds_ResA_ASV,physeq_Andigenum)
sigtab_A_ASV = res_A_ASV[which(res_A_ASV$padj < 0.05), ]
#write.csv(sigtab_A_ASV,"sigtab_A_ASV.csv")


#Troubleshooting DESeq2 data:
#OTU_TABLE_Andigenum = physeq_Andigenum@otu_table
#OTU_TABLE_Andigenum= as.data.frame(OTU_TABLE_Andigenum)
#write.csv(OTU_TABLE_Andigenum,"OTU_Andigenum.csv")


#Chiletanum and Response:
physeq_Chiletanum = subset_samples(ps1, Clade == 'C' & Response != "NA" & Soil_inoculation != "Sterile")
physeq_Chiletanum <- tax_glom(physeq_Chiletanum, taxrank = "Genus", NArm = T)
physeq_Chiletanum = subset_samples(physeq_Chiletanum, taxa_sums(physeq_Chiletanum) > 0)
physeq_Chiletanum = subset_samples(physeq_Chiletanum, sample_sums(physeq_Chiletanum) > 0)
#physeq_Chiletanum = filter_taxa(physeq_Chiletanum, function(x) sum(x > 0) > (0.05*length(x)), TRUE)
diagdds_ResC_ASV = geomMeans_DESeq2_Res(physeq_Chiletanum)
res_C_ASV= res_table(diagdds_ResC_ASV,physeq_Chiletanum)
sigtab_C_ASV = res_C_ASV[which(res_C_ASV$padj < 0.05), ]
#write.csv(sigtab_C_ASV,"sigtab_C_ASV.csv")

#Modern and Response:
physeq_Modern = subset_samples(ps1, Clade == 'M' & Response != "NA" & Soil_inoculation != "Sterile")
physeq_Modern <- tax_glom(physeq_Modern, taxrank = "Genus", NArm = F)
physeq_Modern = subset_samples(physeq_Modern, taxa_sums(physeq_Modern) > 0)
physeq_Modern = subset_samples(physeq_Modern, sample_sums(physeq_Modern) > 0)
#physeq_Modern = filter_taxa(physeq_Modern, function(x) sum(x > 0) > (0.05*length(x)), TRUE)
diagdds_ResM_ASV = geomMeans_DESeq2_Res(physeq_Modern)
res_M_ASV= res_table(diagdds_ResM_ASV,physeq_Modern)
sigtab_M_ASV = res_M_ASV[which(res_M_ASV$padj < 0.05), ]
#write.csv(sigtab_M_ASV,"sigtab_M_ASV.csv")


```


**FIGURE 6 Ch 3 paper**
```{r}
#installation guide here: https://github.com/js229/Vennerable
library(Vennerable)

restab_A_ResP = subset(res_A_ASV, log2FoldChange > 0, Family != "NA")
restab_C_ResP = subset(res_C_ASV, log2FoldChange > 0, Family != "NA")
restab_M_ResP = subset(res_M_ASV, log2FoldChange > 0, Family != "NA")


#generate a list of significant ASVs that are were enriched
my_list <- list(restab_A_ResP$Family,restab_C_ResP$Family,restab_M_ResP$Family)
names(my_list) = c("Andigenum","Chiletanum","Modern")
Vstem <- Venn(my_list)
plot(Vstem, doWeights = FALSE)


```

```{r}
Vstem@IntersectionSets
```

```{r}

restab_A_ResN = subset(res_A_ASV, log2FoldChange < 0, Family != "NA")
restab_C_ResN = subset(res_C_ASV, log2FoldChange < 0, Family != "NA")
restab_M_ResN = subset(res_M_ASV, log2FoldChange < 0, Family != "NA")


my_list <- list(restab_A_ResN$Phylum,restab_C_ResN$Phylum,restab_M_ResN$Phylum)
names(my_list) = c("Andigenum","Chiletanum","Modern")
Vstem <- Venn(my_list)
plot(Vstem, doWeights = FALSE)

Vstem@IntersectionSets

```


#Negatively Selected:
```{r}
#installation guide here: https://github.com/js229/Vennerable
library(Vennerable)

res_A_ASV_ResN = subset(res_A_ASV, log2FoldChange < 0)
res_C_ASV_ResN = subset(res_C_ASV, log2FoldChange < 0)
res_M_ASV_ResN = subset(res_M_ASV, log2FoldChange < 0)


#generate a list of significant ASVs that are were enriched
my_list <- list(res_A_ASV_ResN$Family,sigtab_C_ResN$Family,sigtab_M_ResN$Family)
names(my_list) = c("Andigenum","Chiletanum","Modern")
Vstem <- Venn(my_list)
plot(Vstem, doWeights = FALSE)

Vstem@IntersectionSets

```
```{r}
my_list <- list(sigtab_A_ResN$Phylum,sigtab_C_ResN$Phylum,sigtab_M_ResN$Phylum)
names(my_list) = c("Andigenum","Chiletanum","Modern")
Vstem <- Venn(my_list)
plot(Vstem, doWeights = FALSE)

Vstem@IntersectionSets

```
###Plotting differential abudance of ASVs catagorized as Families and Phyla:

**USE FOR FIGURE 5A**
```{r}
library(RColorBrewer)

TaxOrder = res_A_ASV %>%
  group_by(Phylum)%>%
  summarize(Total=n()) %>%
  arrange(-Total)
TaxOrder = TaxOrder$Phylum
res_A_ASV$Phylum = ordered(res_A_ASV$Phylum, levels=TaxOrder)

nb.cols <- 41 #Number of colors to use for graph for RColorBrewer
palette_colors = colorRampPalette(brewer.pal(15,"Set2"))(nb.cols)

A = ggplot(res_A_ASV,aes(x=Phylum, y=log2FoldChange, alpha= if_else(padj < 0.05,20,5), fill=Phylum))+
  geom_jitter(shape=21, stroke=1, aes(size=baseMean), width=0.2)+
  scale_size(guide=FALSE)+
  scale_fill_manual(values = palette_colors)+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=5))+
  theme(axis.title.y = element_text(size=10))+
  theme(legend.position="none")+
  geom_hline(yintercept=0)+
  ylab(expression(Response~to~Plant~Growth(log[2]-fold~change)))+
  theme(axis.text.x = element_text(face="italic"),
             legend.text = element_text(face="italic"),
             axis.title.x = element_blank())
A

```

**USE FOR FIGURE 5B**

```{r}
TaxOrder = res_C_ASV %>%
  group_by(Phylum)%>%
  summarize(Total=n()) %>%
  arrange(-Total)
TaxOrder = TaxOrder$Phylum
res_C_ASV$Phylum = ordered(res_C_ASV$Phylum, levels=TaxOrder)

nb.cols <- 41 #Number of colors to use for graph for RColorBrewer
palette_colors = colorRampPalette(brewer.pal(15,"Set2"))(nb.cols)

C_ASV_Plot = ggplot(res_C_ASV,aes(x=Phylum, y=log2FoldChange, alpha= if_else(padj < 0.05,20,5), fill=Phylum))+
  geom_jitter(shape=21, stroke=1, aes(size=baseMean), width=0.2)+
  scale_size(guide=FALSE)+
  scale_fill_manual(values = palette_colors)+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=5))+
  theme(axis.title.y = element_text(size=10))+
  theme(legend.position="none")+
  geom_hline(yintercept=0)+
  ylab(expression(Response~to~Plant~Growth(log[2]-fold~change)))+
  theme(axis.text.x = element_text(face="italic"),
             legend.text = element_text(face="italic"),
             axis.title.x = element_blank())
C_ASV_Plot
```


```{r}
TaxOrder = res_M_ASV %>%
  group_by(Phylum)%>%
  summarize(Total=n()) %>%
  arrange(-Total)
TaxOrder = TaxOrder$Phylum
res_M_ASV$Phylum = ordered(res_M_ASV$Phylum, levels=TaxOrder)

nb.cols <- 41 #Number of colors to use for graph for RColorBrewer
palette_colors = colorRampPalette(brewer.pal(15,"Set2"))(nb.cols)

M_ASV_Plot = ggplot(res_M_ASV,aes(x=Phylum, y=log2FoldChange, alpha= if_else(padj < 0.05,20,5), fill=Phylum))+
  geom_jitter(shape=21, stroke=1, aes(size=baseMean), width=0.2)+
  scale_size(guide=FALSE)+
  scale_fill_manual(values = palette_colors)+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=5))+
  theme(axis.title.y = element_text(size=10))+
  theme(legend.position="none")+
  geom_hline(yintercept=0)+
  ylab(expression(Response~to~Plant~Growth(log[2]-fold~change)))+
  theme(axis.text.x = element_text(face="italic"),
             legend.text = element_text(face="italic"),
             axis.title.x = element_blank())
M_ASV_Plot
```


```{r}
#Venn Diagram
sigtab_A_ResP = subset(sigtab_A_ASV, log2FoldChange > 0)
sigtab_C_ResP = subset(sigtab_C_ASV, log2FoldChange > 0)
sigtab_M_ResP = subset(sigtab_M_ASV, log2FoldChange > 0)


#generate a list of significant ASVs that are were enriched
my_list <- list(sigtab_A_ResP$Genus,sigtab_C_ResP$Genus,sigtab_M_ResP$Genus)
names(my_list) = c("Andigenum","Chiletanum","Modern")
Vstem <- Venn(my_list)
plot(Vstem, doWeights = FALSE)
```



###ASV level at iteration 5:


```{r ASV_i5}
#Andigenum and Microbial Response:
physeq_Andigenum = subset_samples(physeq_trimmed, Clade == 'A' & Response != "NA" & Soil_inoculation != "Sterile" & Iteration == 5)
physeq_Andigenum = filter_taxa(physeq_Andigenum, function(x) sum(x > 0) > (0.1*length(x)), TRUE)
diagdds_ResA_ASV = geomMeans_DESeq2_Res(physeq_Andigenum)
res_A_ASV_i5= res_table(diagdds_ResA_ASV,physeq_Andigenum)
sigtab_A_ASV_i5 = res_A_ASV[which(res_A_ASV$padj < 0.05), ]


#Chiletanum and Microbial Response:
physeq_Chiletanum = subset_samples(physeq_trimmed, Clade == 'C' & Response != "NA" & Soil_inoculation != "Sterile" & Iteration == 5)
physeq_Chiletanum = filter_taxa(physeq_Chiletanum, function(x) sum(x > 0) > (0.1*length(x)), TRUE)
diagdds_ResC_ASV = geomMeans_DESeq2_Res(physeq_Chiletanum)
res_C_ASV_i5= res_table(diagdds_ResC_ASV,physeq_Chiletanum)
sigtab_C_ASV_i5 = res_C_ASV[which(res_C_ASV$padj < 0.05), ]
  
#Modern and Microbial Response:
physeq_Modern = subset_samples(physeq_trimmed, Clade == 'M' & Response != "NA" & Soil_inoculation != "Sterile" & Iteration == 5)
physeq_Andigenum = filter_taxa(physeq_Modern, function(x) sum(x > 0) > (0.1*length(x)), TRUE)
diagdds_ResM_ASV = geomMeans_DESeq2_Res(physeq_Modern)
res_M_ASV_i5= res_table(diagdds_ResM_ASV,physeq_Modern)
sigtab_M_ASV_i5 = res_M_ASV[which(res_M_ASV$padj < 0.05), ]

```

```{r}
#Venn Diagram
sigtab_A_ResP_i5 = subset(sigtab_A_ASV_i5, log2FoldChange > 0)
sigtab_C_ResP_i5 = subset(sigtab_C_ASV_i5, log2FoldChange > 0)
sigtab_M_ResP_i5 = subset(sigtab_M_ASV_i5, log2FoldChange > 0)


#generate a list of significant ASVs that are were enriched
my_list <- list(sigtab_A_ResP_i5$Genus,sigtab_C_ResP_i5$Genus,sigtab_M_ResP_i5$Genus)
names(my_list) = c("Andigenum","Chiletanum","Modern")
Vstem <- Venn(my_list)
plot(Vstem, doWeights = FALSE)

my_list <- list(sigtab_A_ResP_i5$Phylum,sigtab_C_ResP_i5$Phylum,sigtab_M_ResP_i5$Phylum)
names(my_list) = c("Andigenum","Chiletanum","Modern")
Vstem <- Venn(my_list)
plot(Vstem, doWeights = FALSE)
```


```{r}
#Venn Diagram
sigtab_A_ResN_i5 = subset(sigtab_A_ASV_i5, log2FoldChange < 0)
sigtab_C_ResN_i5 = subset(sigtab_C_ASV_i5, log2FoldChange < 0)
sigtab_M_ResN_i5 = subset(sigtab_M_ASV_i5, log2FoldChange < 0)


#generate a list of significant ASVs that are were enriched
my_list <- list(sigtab_A_ResN_i5$Family,sigtab_A_ResN_i5$Family,sigtab_A_ResN_i5$Family)
names(my_list) = c("Andigenum","Chiletanum","Modern")
Vstem <- Venn(my_list)
plot(Vstem, doWeights = FALSE)

my_list <- list(sigtab_A_ResN_i5$Phylum,sigtab_A_ResN_i5$Phylum,sigtab_A_ResN_i5$Phylum)
names(my_list) = c("Andigenum","Chiletanum","Modern")
Vstem <- Venn(my_list)
plot(Vstem, doWeights = FALSE)
```
  

### Microbial taxa at Phyla level 


####Co-correlated with Plant response (Across all iterations):

```{r}
#Andigenum growth and Microbial Response:
physeq_Andigenum = subset_samples(ps1, Clade == 'A' & Response != "NA" & Soil_inoculation != "Sterile")
physeq_FamA = tax_glom(physeq_Andigenum,taxrank=rank_names(physeq_Andigenum)[5])
diagdds_ResA_Fam = geomMeans_DESeq2_Res(physeq_FamA)
sigtab_ResA_Fam = DESeq_table(diagdds_ResA_Fam,physeq_FamA)
res_ResA_Fam = res_table(diagdds_ResA_Fam,physeq_FamA)

#Chiletanum growth and Microbial Response
physeq_Chiletanum = subset_samples(ps1, Clade == 'C' & Response != "NA" & Soil_inoculation != "Sterile")
physeq_FamC = tax_glom(physeq_Chiletanum,taxrank=rank_names(physeq_Chiletanum)[5])
#physeq_FamC = filter_taxa(physeq_FamC, function(x) sum(x > 0) > (0.1*length(x)), TRUE)
diagdds_ResC_Fam = geomMeans_DESeq2_Res(physeq_FamC)
res_ResC_Fam = res_table(diagdds_ResC_Fam,physeq_FamC)
sigtab_ResC_Fam = res_ResC_Fam[which(res_ResC_Fam$padj < 0.05), ]


#Modern growth and Microbial Response
physeq_Modern = subset_samples(ps1, Clade == 'M' & Response != "NA" & Soil_inoculation != "Sterile")
physeq_FamM = tax_glom(physeq_Modern,taxrank=rank_names(physeq_Modern)[5])
#physeq_FamM  = filter_taxa(physeq_FamM , function(x) sum(x > 0) > (0.1*length(x)), TRUE)
diagdds_ResM_Fam = geomMeans_DESeq2_Res(physeq_FamM)
res_ResM_Fam = res_table(diagdds_ResM_Fam,physeq_FamM)
sigtab_ResM_Fam = res_ResM_Fam[which(res_ResM_Fam$padj < 0.05), ]
#write.csv(sigtab_ResM_Fam,"sigtab_ResM_Fam.csv")

```


####co-correlated with plant response (iteration 5 only)** 
```{r}
#Andigenum and Microbial Response:
physeq_Andigenum = subset_samples(physeq_trimmed, Clade == 'A' & Response != "NA" & Soil_inoculation != "Sterile" & Iteration == 5)
physeq_FamA = tax_glom(physeq_Andigenum,taxrank=rank_names(physeq_Andigenum)[5])
physeq_FamA = subset_taxa(physeq_FamA, Phylum != "NA")
physeq_FamA = subset_samples(physeq_FamA, sample_sums(physeq_FamA) != 0)
diagdds_ResA_Fam = geomMeans_DESeq2_Res(physeq_FamA)
res_ResA_Fam = res_table(diagdds_ResA_Fam,physeq_FamA)
res_ResA_Fam_i5 = subset(res_ResA_Fam, log2FoldChange != "NA")
sigtab_ResA_Fam_i5 = res_ResA_Fam[which(res_ResA_Fam$padj < 0.05), ]


#Chiletanum and Microbial Response:
physeq_Chiletanum = subset_samples(physeq_trimmed, Clade == 'C' & Response != "NA" & Soil_inoculation != "Sterile" & Iteration == 5)
physeq_FamC = tax_glom(physeq_Chiletanum,taxrank=rank_names(physeq_Chiletanum)[5])
physeq_FamC = subset_taxa(physeq_FamC, Phylum != "NA")
physeq_FamC = subset_samples(physeq_FamC, sample_sums(physeq_FamC) != 0)
diagdds_ResC_Fam = geomMeans_DESeq2_Res(physeq_FamC)
res_ResC_Fam = res_table(diagdds_ResC_Fam,physeq_FamC)
res_ResC_Fam_i5 = subset(res_ResC_Fam, log2FoldChange != "NA")
sigtab_ResC_Fam_i5 = res_ResC_Fam[which(res_ResC_Fam$padj < 0.05), ]
  
#Modern and Microbial Response:
physeq_Modern = subset_samples(physeq_trimmed, Clade == 'M' & Response != "NA" & Soil_inoculation != "Sterile" & Iteration == 5)
physeq_FamM = tax_glom(physeq_Modern,taxrank=rank_names(physeq_Modern)[5])
physeq_FamM = subset_taxa(physeq_FamM, Family != "NA")
physeq_FamM = subset_samples(physeq_FamM, sample_sums(physeq_FamM) != 0)
diagdds_ResM_Fam = geomMeans_DESeq2_Res(physeq_FamM)
res_ResM_Fam = res_table(diagdds_ResM_Fam,physeq_FamM)
res_ResM_Fam_i5 = subset(res_ResM_Fam, log2FoldChange != "NA")
sigtab_ResM_Fam_i5 = res_ResM_Fam[which(res_ResM_Fam$padj < 0.05), ]

```


```{r}
sigtab_ResA_FamP = subset(sigtab_ResA_Fam_i5, log2FoldChange > 0)
sigtab_ResC_FamP = subset(sigtab_ResC_Fam_i5, log2FoldChange > 0)
sigtab_ResM_FamP = subset(sigtab_ResM_Fam_i5, log2FoldChange > 0)

my_list <- list(sigtab_ResC_FamP$Family,sigtab_ResC_FamP$Family,sigtab_ResM_FamP$Family)
names(my_list) = c("Andigenum","Chiletanum","Modern")
Vstem <- Venn(my_list)
plot(Vstem, doWeights = FALSE)

Vstem@IntersectionSets
```


Results: **Iteration 5 only**

Andigenum: Saccharimonadales (Positive), Corynebacteriaceae (negative), Cytophagaceae (Positive), Crocinitomicaceae (Positive), Microscillaceae (Positive), Pseudomonadaceae (negative), Propionibacteriaceae (negative), Haliangiaceae (Positive), Flavobacteriaceae (Negative)

Chiletanum: None

Modern: Cyclobacteriaceae (Positive), PeM15 (Actinobacteria), Puniceispirillales (Positive), Spirosomaceae (Positive)



```{r}
TaxOrder = res_ResA_Fam %>%
  group_by(Family)%>%
  summarize(Total=n()) %>%
  arrange(-Total)
TaxOrder = TaxOrder$Family
res_ResA_Fam$Family = ordered(res_ResA_Fam$Family, levels=TaxOrder)

nb.cols <- 109 #Number of colors to use for graph for RColorBrewer
palette_colors = colorRampPalette(brewer.pal(8,"Set2"))(nb.cols)

A_Family_Plot = ggplot(res_ResA_Fam,aes(x=Family, y=log2FoldChange, alpha= if_else(padj < 0.05,20,5), fill=Phylum))+
  geom_jitter(shape=21, stroke=1, aes(size=baseMean), width=0.2)+
  scale_size(guide=FALSE)+
  scale_fill_manual(values = palette_colors)+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=5))+
  theme(axis.title.y = element_text(size=10))+
  theme(legend.position="none")+
  geom_hline(yintercept=0)+
  ylab(expression(Response~to~Plant~Growth(log[2]-fold~change)))+
  theme(axis.text.x = element_text(face="italic"),
             legend.text = element_text(face="italic"),
             axis.title.x = element_blank())
A_Family_Plot

```

```{r}
TaxOrder = res_ResC_Fam %>%
  group_by(Phylum)%>%
  summarize(Total=n()) %>%
  arrange(-Total)
TaxOrder = TaxOrder$Phylum
res_ResC_Fam$Phylum = ordered(res_ResC_Fam$Phylum, levels=TaxOrder)

nb.cols <- 41 #Number of colors to use for graph for RColorBrewer
palette_colors = colorRampPalette(brewer.pal(8,"Set2"))(nb.cols)

C_Family_Plot = ggplot(res_ResC_Fam,aes(x=Phylum, y=log2FoldChange, alpha= if_else(padj < 0.05,20,5), fill=Phylum))+
  geom_jitter(shape=21, stroke=1, aes(size=baseMean), width=0.2)+
  scale_size(guide=FALSE)+
  scale_fill_manual(values = palette_colors)+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=5))+
  theme(axis.title.y = element_text(size=10))+
  theme(legend.position="none")+
  geom_hline(yintercept=0)+
  ylab(expression(Response~to~Plant~Growth(log[2]-fold~change)))+
  theme(axis.text.x = element_text(face="italic"),
             legend.text = element_text(face="italic"),
             axis.title.x = element_blank())
C_Family_Plot
```


```{r}
TaxOrder = res_ResM_Fam %>%
  group_by(Phylum)%>%
  summarize(Total=n()) %>%
  arrange(-Total)
TaxOrder = TaxOrder$Phylum
res_ResM_Fam$Phylum = ordered(res_ResM_Fam$Phylum, levels=TaxOrder)

nb.cols <- 41 #Number of colors to use for graph for RColorBrewer
palette_colors = colorRampPalette(brewer.pal(8,"Set2"))(nb.cols)

M_Family_Plot = ggplot(res_ResM_Fam,aes(x=Phylum, y=log2FoldChange, alpha= if_else(padj < 0.05,20,5), fill=Phylum))+
  geom_jitter(shape=21, stroke=1, aes(size=baseMean), width=0.2)+
  scale_size(guide=FALSE)+
  scale_fill_manual(values = palette_colors)+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=5))+
  theme(axis.title.y = element_text(size=10))+
  theme(legend.position="none")+
  geom_hline(yintercept=0)+
  ylab(expression(Response~to~Plant~Growth(log[2]-fold~change)))+
  theme(axis.text.x = element_text(face="italic"),
             legend.text = element_text(face="italic"),
             axis.title.x = element_blank())
M_Family_Plot
```



###At Phylum Level: 
```{r}
#Andigenum growth and Microbial Response:
physeq_Andigenum = subset_samples(physeq_trimmed, Clade == 'A' & Response != "NA" & Soil_inoculation != "Sterile")
physeq_phylaA = tax_glom(physeq_Andigenum,taxrank=rank_names(physeq_Andigenum)[2])
physeq_phylaA = subset_taxa(physeq_phylaA, Phylum != "NA")
physeq_phylaA = subset_samples(physeq_phylaA, sample_sums(physeq_phylaA) != 0)
diagdds_ResA_Phyla = geomMeans_DESeq2_Res(physeq_phylaA)
sigtab_ResA_Phyla = DESeq_table(diagdds_ResA_Phyla,physeq_phylaA)
res_ResA_Phyla = res_table(diagdds_ResA_Phyla,physeq_phylaA)
sigtab_ResA_Phyla = res_ResA_Phyla[which(res_ResA_Phyla$padj < 0.05), ]
#write.csv(sigtab_ResA_Phyla,"sigtab_ResA_Phyla.csv")

#Chiletanum growth and Microbial Response
physeq_Chiletanum = subset_samples(physeq_trimmed, Clade == 'C' & Response != "NA" & Soil_inoculation != "Sterile")
physeq_phylaC = tax_glom(physeq_Chiletanum,taxrank=rank_names(physeq_Chiletanum)[2])
physeq_phylaC = subset_taxa(physeq_phylaC, Phylum != "NA")
physeq_phylaC = subset_samples(physeq_phylaC, sample_sums(physeq_phylaC) != 0)
diagdds_ResC_Phyla = geomMeans_DESeq2_Res(physeq_phylaC)
res_ResC_Phyla = res_table(diagdds_ResC_Phyla,physeq_phylaC)
sigtab_ResC_Phyla = res_ResC_Phyla[which(res_ResC_Phyla$padj < 0.05), ]


#Modern growth and Microbial Response
physeq_Modern = subset_samples(physeq_trimmed, Clade == 'M' & Response != "NA" & Soil_inoculation != "Sterile")
physeq_phylaM = tax_glom(physeq_Modern,taxrank=rank_names(physeq_Modern)[2])
physeq_phylaM = subset_taxa(physeq_phylaM, Phylum != "NA")
physeq_phylaM = subset_samples(physeq_phylaM, sample_sums(physeq_phylaM) != 0)
diagdds_ResM_Phyla = geomMeans_DESeq2_Res(physeq_phylaM)
res_ResM_Phyla = res_table(diagdds_ResM_Phyla,physeq_phylaM)
sigtab_ResM_Phyla = res_ResM_Phyla[which(res_ResM_Phyla$padj < 0.05), ]
#write.csv(sigtab_ResM_Phyla,"sigtab_ResM_Phyla.csv")

```


####PLOTS for Phyla differences


**Andigenum**
```{r}
library(RColorBrewer)

TaxOrder = res_ResA_Phyla %>%
  group_by(Phylum)%>%
  summarize(Total=n()) %>%
  arrange(-Total)
TaxOrder = TaxOrder$Phylum
res_ResA_Phyla$Phylum = ordered(res_ResA_Phyla$Phylum, levels=TaxOrder)

nb.cols <- 41 #Number of colors to use for graph for RColorBrewer
palette_colors = colorRampPalette(brewer.pal(8,"Set2"))(nb.cols)

A_Phyla_Plot = ggplot(res_ResA_Phyla,aes(x=Phylum, y=log2FoldChange, alpha= if_else(padj < 0.05,20,5), fill=Phylum))+
  geom_jitter(shape=21, stroke=1, aes(size=baseMean), width=0.2)+
  scale_size(guide=FALSE)+
  scale_fill_manual(values = palette_colors)+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=5))+
  theme(axis.title.y = element_text(size=10))+
  theme(legend.position="none")+
  geom_hline(yintercept=0)+
  ylab(expression(Response~to~Plant~Growth(log[2]-fold~change)))+
  theme(axis.text.x = element_text(face="italic"),
             legend.text = element_text(face="italic"),
             axis.title.x = element_blank())
A_Phyla_Plot
```


```{r}
TaxOrder = res_ResC_Phyla %>%
  group_by(Phylum)%>%
  summarize(Total=n()) %>%
  arrange(-Total)
TaxOrder = TaxOrder$Phylum
res_ResC_Phyla$Phylum = ordered(res_ResC_Phyla$Phylum, levels=TaxOrder)

nb.cols <- 41 #Number of colors to use for graph for RColorBrewer
palette_colors = colorRampPalette(brewer.pal(8,"Set2"))(nb.cols)

C_Phyla_Plot = ggplot(res_ResC_Phyla,aes(x=Phylum, y=log2FoldChange, alpha= if_else(padj < 0.05,20,5), fill=Phylum))+
  geom_jitter(shape=21, stroke=1, aes(size=baseMean), width=0.2)+
  scale_size(guide=FALSE)+
  scale_fill_manual(values = palette_colors)+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=5))+
  theme(axis.title.y = element_text(size=10))+
  theme(legend.position="none")+
  geom_hline(yintercept=0)+
  ylab(expression(Response~to~Plant~Growth(log[2]-fold~change)))+
  theme(axis.text.x = element_text(face="italic"),
             legend.text = element_text(face="italic"),
             axis.title.x = element_blank())
C_Phyla_Plot
```



```{r}
TaxOrder = res_ResM_Phyla %>%
  group_by(Phylum)%>%
  summarize(Total=n()) %>%
  arrange(-Total)
TaxOrder = TaxOrder$Phylum
res_ResM_Phyla$Phylum = ordered(res_ResM_Phyla$Phylum, levels=TaxOrder)

nb.cols <- 41 #Number of colors to use for graph for RColorBrewer
palette_colors = colorRampPalette(brewer.pal(8,"Set2"))(nb.cols)

M_Phyla_Plot = ggplot(res_ResM_Phyla,aes(x=Phylum, y=log2FoldChange, alpha= if_else(padj < 0.05,20,5), fill=Phylum))+
  geom_jitter(shape=21, stroke=1, aes(size=baseMean), width=0.2)+
  scale_size(guide=FALSE)+
  scale_fill_manual(values = palette_colors)+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=5))+
  theme(axis.title.y = element_text(size=10))+
  theme(legend.position="none")+
  geom_hline(yintercept=0)+
  ylab(expression(Response~to~Plant~Growth(log[2]-fold~change)))+
  theme(axis.text.x = element_text(face="italic"),
             legend.text = element_text(face="italic"),
             axis.title.x = element_blank())
M_Phyla_Plot
```



**Results**: for microbial phyla abundance and plant response across time:

Andigenum: Patescibacteria, Dependentiae both are positive
Chiletanum: None
Modern: None

###Microbial taxa (Phyla) differentially abundant and Plant response at iteration 5
```{r}
#Andigenum growth and Microbial Response:
physeq_Andigenum = subset_samples(physeq_trimmed, Clade == 'A' & Response != "NA" & Soil_inoculation != "Sterile" & Iteration == 5)
physeq_phylaA = tax_glom(physeq_Andigenum,taxrank=rank_names(physeq_Andigenum)[2])
physeq_phylaA = subset_taxa(physeq_phylaA, Phylum != "NA")
physeq_phylaA = subset_samples(physeq_phylaA, sample_sums(physeq_phylaA) != 0)
diagdds_ResA_Phyla = geomMeans_DESeq2_Res(physeq_phylaA)
sigtab_ResA_Phyla = DESeq_table(diagdds_ResA_Phyla,physeq_phylaA)
res_ResA_Phyla = res_table(diagdds_ResA_Phyla,physeq_phylaA)

#Chiletanum growth and Microbial Response
physeq_Chiletanum = subset_samples(physeq_trimmed, Clade == 'C' & Response != "NA" & Soil_inoculation != "Sterile" & Iteration == 5)
physeq_phylaC = tax_glom(physeq_Chiletanum,taxrank=rank_names(physeq_Chiletanum)[2])
physeq_phylaC = subset_taxa(physeq_phylaC, Phylum != "NA")
physeq_phylaC = subset_samples(physeq_phylaC, sample_sums(physeq_phylaC) != 0)
diagdds_ResC_Phyla = geomMeans_DESeq2_Res(physeq_phylaC)
res_ResC_Phyla = res_table(diagdds_ResC_Phyla,physeq_phylaC)
sigtab_ResC_Phyla = res_ResC_Phyla[which(res_ResC_Phyla$padj < 0.05), ]


#Modern growth and Microbial Response
physeq_Modern = subset_samples(physeq_trimmed, Clade == 'M' & Response != "NA" & Soil_inoculation != "Sterile" & Iteration == 5)
physeq_phylaM = tax_glom(physeq_Modern,taxrank=rank_names(physeq_Modern)[2])
physeq_phylaM = subset_taxa(physeq_phylaM, Phylum != "NA")
physeq_phylaM = subset_samples(physeq_phylaM, sample_sums(physeq_phylaM) != 0)
diagdds_ResM_Phyla = geomMeans_DESeq2_Res(physeq_phylaM)
res_ResM_Phyla = res_table(diagdds_ResM_Phyla,physeq_phylaM)
sigtab_ResM_Phyla = res_ResM_Phyla[which(res_ResM_Phyla$padj < 0.05), ]

```
Results: 

Similar to what was previously seen for Andigneum: Patescibacteria, Dependentiae

Patescibacteria --> https://www.frontiersin.org/articles/10.3389/fmicb.2020.01948/full

Discussion: Phyla potentially generalist that have a positive impact on plant growth...


##Let's examine DESeq2 for Response*iteration for each Clade: 

**Note** : Looking for consistent Deseq ~ response x iteration. Need to think what the positive value means vs negative value mean….baseline iteration 1 at zero…Taxa into camps interaction terms are significant vs consistent

This should teach us about interaction between taxa and plant. Ought to be stable. If pathogen, negative…. Commensals or mutualist. Relationship could be different amongst Clade. Venn Diagram....


```{r}
#Function 5:
geomMeans_DESeq2_ResXIt = function(ps_object){
diagdds_clade= phyloseq_to_deseq2(ps_object, ~ Response * Iteration)

gm_mean = function(x, na.rm=TRUE){
  exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
}

geoMeans = apply(counts(diagdds_clade), 1, gm_mean)
diagdds = estimateSizeFactors(diagdds_clade, geoMeans = geoMeans)

diagdds_clade = DESeq(diagdds, fitType="parametric")
return(diagdds_clade)
}
```


###At ASV level:

Response * Iteration : Microbes that have been there consistently...with positive plant response....

```{r}
#Andigenum and Response over time:
physeq_Andigenum = subset_samples(physeq_trimmed, Clade == 'A' & Response != "NA" & Soil_inoculation != "Sterile")
physeq_Andigenum = subset_samples(physeq_Andigenum, sample_sums(physeq_Andigenum) != 0)
diagdds_ResXItA_ASV = geomMeans_DESeq2_ResXIt(physeq_Andigenum)
res_ResXItA_ASV= res_table(diagdds_ResXItA_ASV,physeq_Andigenum)
sigtab_ResXItA_ASV = res_ResXItA_ASV[which(res_ResXItA_ASV$padj < 0.05), ]


#Chiletanum and Response over time:
physeq_Chiletanum = subset_samples(physeq_trimmed, Clade == 'C' & Response != "NA" & Soil_inoculation != "Sterile")
physeq_Chiletanum = subset_samples(physeq_Chiletanum, sample_sums(physeq_Chiletanum) != 0)
diagdds_ResXItC_ASV = geomMeans_DESeq2_ResXIt(physeq_Chiletanum)
res_ResXItC_ASV= res_table(diagdds_ResXItC_ASV,physeq_Chiletanum)
sigtab_ResXItC_ASV = res_ResXItC_ASV[which(res_ResXItC_ASV$padj < 0.05), ]


#Modern and Response over time:
physeq_Modern = subset_samples(physeq_trimmed, Clade == 'M' & Response != "NA" & Soil_inoculation != "Sterile")
physeq_Modern = subset_samples(physeq_Modern, sample_sums(physeq_Modern) != 0)
diagdds_ResXItM_ASV = geomMeans_DESeq2_ResXIt(physeq_Modern)
res_ResXItM_ASV= res_table(diagdds_ResXItM_ASV,physeq_Modern)
sigtab_ResXItM_ASV = res_ResXItM_ASV[which(res_ResXItM_ASV$padj < 0.05), ]

```
At ASV level: We do see a lot more differentially abundant ASVs for moderns... 

```{r}
sigtab_ResA_FamP = subset(sigtab_ResXItA_ASV, log2FoldChange > 0)
sigtab_ResC_FamP = subset(sigtab_ResXItC_ASV, log2FoldChange > 0)
sigtab_ResM_FamP = subset(sigtab_ResXItM_ASV, log2FoldChange > 0)

my_list <- list(sigtab_ResC_FamP$Family,sigtab_ResC_FamP$Family,sigtab_ResM_FamP$Family)
names(my_list) = c("Andigenum","Chiletanum","Modern")
Vstem <- Venn(my_list)
plot(Vstem, doWeights = FALSE)

Vstem@IntersectionSets
```

```{r}
sigtab_ResA_FamN = subset(sigtab_ResXItA_ASV, log2FoldChange < 0)
sigtab_ResC_FamN = subset(sigtab_ResXItC_ASV, log2FoldChange < 0)
sigtab_ResM_FamN = subset(sigtab_ResXItM_ASV, log2FoldChange < 0)

my_list <- list(sigtab_ResC_FamN$Family,sigtab_ResC_FamN$Family,sigtab_ResM_FamN$Family)
names(my_list) = c("Andigenum","Chiletanum","Modern")
Vstem <- Venn(my_list)
plot(Vstem, doWeights = FALSE)

Vstem@IntersectionSets
```


***

#Plant growth and Microbial Response with Time (~ Response * Time):

##At Phyla Level:
```{r}

physeq_Andigenum = subset_samples(physeq_trimmed, Clade == 'A' & Response != "NA" & Soil_inoculation != "Sterile")
physeq_phylaA = tax_glom(physeq_Andigenum,taxrank=rank_names(physeq_Andigenum)[2])
physeq_phylaA = subset_taxa(physeq_phylaA, Phylum != "NA")
physeq_phylaA = subset_samples(physeq_phylaA, sample_sums(physeq_phylaA) != 0)
diagdds_ResA_Phyla = geomMeans_DESeq2_ResXIt(physeq_phylaA)
sigtab_ResA_PhyTime = DESeq_table(diagdds_ResA_Phyla,physeq_phylaA)
res_A_PhyTime = res_table(diagdds_ResA_Phyla,physeq_phylaA)

#Chiletanum growth and Microbial Response
physeq_Chiletanum = subset_samples(physeq_trimmed, Clade == 'C' & Response != "NA" & Soil_inoculation != "Sterile")
physeq_phylaC = tax_glom(physeq_Chiletanum,taxrank=rank_names(physeq_Chiletanum)[2])
physeq_phylaC = subset_taxa(physeq_phylaC, Phylum != "NA")
physeq_phylaC = subset_samples(physeq_phylaC, sample_sums(physeq_phylaC) != 0)
diagdds_ResC_Phyla = geomMeans_DESeq2_ResXIt(physeq_phylaC)
res_C_PhyTime = res_table(diagdds_ResC_Phyla,physeq_phylaC)
sigtab_C_PhyTime = res_C_PhyTime[which(res_C_PhyTime$padj < 0.05), ]


#Modern growth and Microbial Response
physeq_Modern = subset_samples(physeq_trimmed, Clade == 'M' & Response != "NA" & Soil_inoculation != "Sterile")
physeq_phylaM = tax_glom(physeq_Modern,taxrank=rank_names(physeq_Modern)[2])
physeq_phylaM = subset_taxa(physeq_phylaM, Phylum != "NA")
physeq_phylaM = subset_samples(physeq_phylaM, sample_sums(physeq_phylaM) != 0)
diagdds_ResM_Phyla = geomMeans_DESeq2_ResXIt(physeq_phylaM)
res_M_PhyTime = res_table(diagdds_ResM_Phyla,physeq_phylaM)
sigtab_M_PhyTime = res_M_PhyTime[which(res_M_PhyTime$padj < 0.05), ]

```

```{r}
sigtab_ResA_FamP = subset(sigtab_ResA_Fam, log2FoldChange > 0)
sigtab_ResC_FamP = subset(sigtab_ResC_Fam, log2FoldChange > 0)
sigtab_ResM_FamP = subset(sigtab_ResM_Fam, log2FoldChange > 0)

my_list <- list(sigtab_ResC_FamP$Family,sigtab_ResC_FamP$Family,sigtab_ResM_FamP$Family)
names(my_list) = c("Andigenum","Chiletanum","Modern")
Vstem <- Venn(my_list)
plot(Vstem, doWeights = FALSE)

Vstem@IntersectionSets
```


**Results:** 
Chiletanum: Firmicutes (Positive), Chloroflexi (Negative), Desulfobacterota (Negative), Proteobacteria (Positive), Actinobacteriota (Positive), Acidobacteriota (Negative) with Plant growth with time

Andigneum: Dependentiae Positive with Plant Growth with Time

Moderns: None

##At Family Level:
```{r}
physeq_Andigenum = subset_samples(physeq_trimmed, Clade == 'A' & Response != "NA" & Soil_inoculation != "Sterile")
physeq_FamilyA = tax_glom(physeq_Andigenum,taxrank=rank_names(physeq_Andigenum)[5])
physeq_FamilyA = subset_taxa(physeq_FamilyA, Family != "NA")
physeq_FamilyA = subset_samples(physeq_FamilyA, sample_sums(physeq_FamilyA) != 0)
diagdds_ResA_Family = geomMeans_DESeq2_ResXIt(physeq_FamilyA)
res_ResA_FamTime = res_table(diagdds_ResA_Family,physeq_FamilyA)
sigtab_ResA_FamTime = res_ResA_FamTime[which(res_ResA_FamTime$padj < 0.05), ]

#Chiletanum growth and Microbial Response
physeq_Chiletanum = subset_samples(physeq_trimmed, Clade == 'C' & Response != "NA" & Soil_inoculation != "Sterile")
physeq_FamilyC = tax_glom(physeq_Chiletanum,taxrank=rank_names(physeq_Chiletanum)[5])
physeq_FamilyC = subset_taxa(physeq_FamilyC, Family != "NA")
physeq_FamilyC = subset_samples(physeq_FamilyC, sample_sums(physeq_FamilyC) != 0)
diagdds_ResC_Family = geomMeans_DESeq2_ResXIt(physeq_FamilyC)
res_ResC_FamTime = res_table(diagdds_ResC_Family,physeq_FamilyC)
sigtab_ResC_FamTime = res_ResC_Family[which(res_ResC_Family$padj < 0.05), ]


#Modern growth and Microbial Response
physeq_Modern = subset_samples(physeq_trimmed, Clade == 'M' & Response != "NA" & Soil_inoculation != "Sterile")
physeq_FamilyM = tax_glom(physeq_Modern,taxrank=rank_names(physeq_Modern)[5])
physeq_FamilyM = subset_taxa(physeq_FamilyM, Phylum != "NA")
physeq_FamilyM = subset_samples(physeq_FamilyM, sample_sums(physeq_FamilyM) != 0)
diagdds_ResM_Family = geomMeans_DESeq2_ResXIt(physeq_FamilyM)
res_ResM_FamTime = res_table(diagdds_ResM_Family ,physeq_FamilyM)
sigtab_ResM_FamTime = res_ResM_Family[which(res_ResM_Family$padj < 0.05), ]

```

**Results:** 
Chiletanum:Xanthomonadales (Order), Bacillaceae(Family), R7C24 (family) all enriched with positive plant growth with time 

Andigeneum: 

Moderns: Flavobacteriales (Order)
