---
title: "DESeq2_Clade"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

```{r}
library(phyloseq)
library(tidyverse)
library(ggplot2)
library(lme4)
library(lmerTest)
library(emmeans)
library(DESeq2)
```


*Read phyloseq file:* 

```{r}
physeq_trimmed = readRDS(file = "ps1.RDS")
```

Filtering based on prevalance of taxa: based on https://github.com/spholmes/F1000_workflow/blob/master/MicrobiomeWorkflow/MicrobiomeWorkflowII.Rmd


#DESeq2 functions 
Getting tired to cutting pasting DESeq2 commands let's write a function for it:

```{r}
#Function 1:
geomMeans_DESeq2 = function(ps_object){
diagdds_clade= phyloseq_to_deseq2(ps_object, ~ Iteration)

gm_mean = function(x, na.rm=TRUE){
  exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
}

geoMeans = apply(counts(diagdds_clade), 1, gm_mean)
diagdds = estimateSizeFactors(diagdds_clade, geoMeans = geoMeans)

diagdds_clade = DESeq(diagdds, fitType="local")
return(diagdds_clade)
}


#Function 1.5:
geomMeans_DESeq2_Clade = function(ps_object){
diagdds_clade= phyloseq_to_deseq2(ps_object, ~ Clade)

gm_mean = function(x, na.rm=TRUE){
  exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
}

geoMeans = apply(counts(diagdds_clade), 1, gm_mean)
diagdds = estimateSizeFactors(diagdds_clade, geoMeans = geoMeans)

diagdds_clade = DESeq(diagdds, fitType="local")
return(diagdds_clade)
}

#Function 2: significant taxa < alpha
DESeq_table = function(DESeq2_ds,ps_object,X){
res = results(DESeq2_ds, cooksCutoff = FALSE)
alpha = X
sigtab = res[which(res$padj < alpha), ]
sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(ps_object)[rownames(sigtab), ], "matrix"))
return(sigtab)
}

#Function 3: All results

res_table = function(diagdds_clade,ps_object){
  res = results(diagdds_clade, cooksCutoff = FALSE)
  res_A_Phyla = cbind(as(res, "data.frame"), as(tax_table(ps_object)[rownames(res), ], "matrix"))
}

```

*testing out our function*

```{r}
physeq_Andigenum = subset_samples(physeq_trimmed, Clade == 'A' & Response != "NA" & Soil_inoculation != "Sterile")
physeq_Andigenum = subset_samples(physeq_Andigenum, sample_sums(physeq_Andigenum) != 0)
diagdds_cladeA_test = geomMeans_DESeq2(physeq_Andigenum)

#Looks like our deomMeans_Deseq2 function works!! based on our test 

sigtab_A_test = DESeq_table(diagdds_cladeA_test,physeq_Andigenum,0.05)

#Looks like our DESeq_table function now works too!

```




```{r}
physeq_Andigenum = subset_samples(ps1, Clade == 'A' & Response != "NA" & Soil_inoculation != "Sterile")
physeq_Andigenum = subset_samples(physeq_Andigenum, sample_sums(physeq_Andigenum) != 0)
diagdds_cladeA = geomMeans_DESeq2(physeq_Andigenum)
sigtab_A = DESeq_table(diagdds_cladeA,physeq_Andigenum,0.05)
res_A_ASV = res_table(diagdds_cladeA,physeq_Andigenum)

#At Phylum Level: 
physeq_phylaA = tax_glom(physeq_Andigenum,taxrank=rank_names(physeq_Andigenum)[2])
physeq_phylaA = subset_taxa(physeq_phylaA, Phylum != "NA")
physeq_phylaA = subset_samples(physeq_phylaA, sample_sums(physeq_phylaA) != 0)

#At Family Level: 
physeq_FamA = tax_glom(physeq_Andigenum,taxrank=rank_names(physeq_Andigenum)[5])
physeq_FamA = subset_taxa(physeq_FamA, Phylum != "NA")
physeq_FamA = subset_samples(physeq_FamA, sample_sums(physeq_FamA) != 0)

#DESeq2 function: 
diagdds_cladeA_Phyla = geomMeans_DESeq2(physeq_phylaA)
diagdds_cladeA_Fam = geomMeans_DESeq2(physeq_FamA)

#Making Tables for significant taxa and all taxas:
sigtab_A_Phyla = DESeq_table(diagdds_cladeA_Phyla,physeq_phylaA,0.05)
sigtab_A_Fam = DESeq_table(diagdds_cladeA_Fam,physeq_FamA,0.05)
res_A_Phyla = res_table(diagdds_cladeA_Phyla,physeq_phylaA)
res_A_Fam = res_table(diagdds_cladeA_Fam,physeq_FamA)
```



###Graphing scatterplot of ASVs 

grouped by family (Not interested)
```{r}
# Ordering factor by phyla in alphabetical order
res_A_ASV = subset(res_A_ASV, Phylum != "NA" & log2FoldChange != "NA" & padj != "NA")
TaxOrder = res_A_ASV %>%
  group_by(Phylum)%>%
  summarize(Total=n()) %>%
  arrange(-Total)
TaxOrder = TaxOrder$Phylum
res_A_ASV$Phylum = ordered(res_A_ASV$Phylum, levels=TaxOrder)

library(RColorBrewer)
nb.cols <- 36 #Number of colors to use for graph for RColorBrewer
palette_colors = colorRampPalette(brewer.pal(8,"Set2"))(nb.cols)

A = ggplot(res_A_ASV,aes(x=Phylum, y=log2FoldChange, alpha= if_else(padj < 0.05,20,5), fill=Phylum))+
  geom_jitter(shape=21, stroke=1, aes(size=baseMean), width=0.2)+
  scale_size(guide=FALSE)+
  scale_fill_manual(values = palette_colors)+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=5))+
  theme(axis.title.y = element_text(size=10))+
  theme(legend.position="none")+
  geom_hline(yintercept=0)+
  ylab(expression(Response~to~Andigenum (log[2]-fold~change)))+
  theme(axis.text.x = element_text(face="italic"),
             legend.text = element_text(face="italic"),
             axis.title.x = element_blank())
A
```


grouped by Phyla **USE**
```{r}
res_A_Phyla = subset(res_A_Phyla, Phylum != "NA" & log2FoldChange != "NA" & padj != "NA")
TaxOrder = res_A_Phyla %>%
  group_by(Phylum)%>%
  summarize(Total=n()) %>%
  arrange(-Total)
TaxOrder = TaxOrder$Phylum
res_A_Phyla$Phylum = ordered(res_A_Phyla$Phylum, levels=TaxOrder)

library(RColorBrewer)
nb.cols <- 36
palette_colors = colorRampPalette(brewer.pal(8,"Set2"))(nb.cols)

A = ggplot(res_A_Phyla,aes(x=Phylum, y=log2FoldChange, alpha= if_else(padj < 0.05,20,5), fill=Phylum))+
  geom_jitter(shape=21, stroke=1, aes(size=baseMean), width=0.2)+
  scale_size(guide=FALSE)+
  scale_fill_manual(values = palette_colors)+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=5))+
  theme(axis.title.y = element_text(size=10))+
  theme(legend.position="none")+
  geom_hline(yintercept=0)+
  ylab(expression(Response~to~Andigenum(log[2]-fold~change)))+
  theme(axis.text.x = element_text(face="italic"),
             legend.text = element_text(face="italic"),
             axis.title.x = element_blank())
A
```


#Chiletanum

```{r}
physeq_Chiletanum = subset_samples(physeq_trimmed, Clade == 'C' & Response != "NA" & Soil_inoculation != "Sterile")
physeq_Chiletanum = subset_samples(physeq_Chiletanum, sample_sums(physeq_Chiletanum) != 0)

diagdds_cladeC = geomMeans_DESeq2(physeq_Chiletanum)
sigtab_C = DESeq_table(diagdds_cladeC,physeq_Chiletanum,0.05)
res_C = res_table(diagdds_cladeC,physeq_Chiletanum)

#At Phylum Level: 
physeq_phylaC = tax_glom(physeq_Chiletanum,taxrank=rank_names(physeq_Chiletanum)[2])
physeq_phylaC = subset_taxa(physeq_phylaC, Phylum != "NA")
physeq_phylaC = subset_samples(physeq_phylaC, sample_sums(physeq_phylaC) != 0)

#At Family Level: 
physeq_FamC = tax_glom(physeq_Chiletanum,taxrank=rank_names(physeq_Chiletanum)[5])
physeq_FamC = subset_taxa(physeq_FamC, Phylum != "NA")
physeq_FamC = subset_samples(physeq_FamC, sample_sums(physeq_FamC) != 0)

#DESeq2 function: 
diagdds_cladeC_Phyla = geomMeans_DESeq2(physeq_phylaC)
diagdds_cladeC_Fam = geomMeans_DESeq2(physeq_FamC)

#Making Tables for significant taxa and all taxas:
sigtab_C_Phyla = DESeq_table(diagdds_cladeC_Phyla,physeq_phylaC,0.05)
sigtab_C_Fam = DESeq_table(diagdds_cladeC_Fam,physeq_FamC,0.05)
res_C_Phyla = res_table(diagdds_cladeC_Phyla,physeq_phylaC)
res_C_Fam = res_table(diagdds_cladeC_Fam,physeq_FamC)
```



###Chiletanum plots (**USE for Ch2 SI Figure 2**)
```{r}
# Ordering factor by most abundant phyla
res_C_Phyla = subset(res_C_Phyla, Phylum != "NA" & log2FoldChange != "NA" & padj != "NA")
TaxOrder = res_C_Phyla %>%
  group_by(Phylum)%>%
  summarize(Total=n()) %>%
  arrange(-Total)
TaxOrder = TaxOrder$Phylum
res_C_Phyla$Phylum = ordered(res_C_Phyla$Phylum, levels=TaxOrder)

library(RColorBrewer)
nb.cols <- 36
palette_colors = colorRampPalette(brewer.pal(8,"Set2"))(nb.cols)

C = ggplot(res_C_Phyla,aes(x=Phylum, y=log2FoldChange, alpha= if_else(padj < 0.05,20,5), fill=Phylum))+
  geom_jitter(shape=21, stroke=1, aes(size=baseMean), width=0.2)+
  scale_size(guide=FALSE)+
  scale_fill_manual(values = palette_colors)+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=5))+
  theme(axis.title.y = element_text(size=10))+
  theme(legend.position="none")+
  geom_hline(yintercept=0)+
  ylab(expression(Response~to~Chiletanum(log[2]-fold~change)))+
  theme(axis.text.x = element_text(face="italic"),
             legend.text = element_text(face="italic"),
             axis.title.x = element_blank())
C
```


```{r}
res_C_ASV = subset(res_C, Phylum != "NA" & log2FoldChange != "NA" & padj != "NA")
TaxOrder = res_C_ASV %>%
  group_by(Phylum)%>%
  summarize(Total=n()) %>%
  arrange(-Total)
TaxOrder = TaxOrder$Phylum
res_C_ASV$Phylum = ordered(res_C_ASV$Phylum, levels=TaxOrder)

library(RColorBrewer)
nb.cols <- 36 #Number of colors to use for graph for RColorBrewer
palette_colors = colorRampPalette(brewer.pal(8,"Set2"))(nb.cols)

C = ggplot(res_C_ASV,aes(x=Phylum, y=log2FoldChange, alpha= if_else(padj < 0.05,20,5), fill=Phylum))+
  geom_jitter(shape=21, stroke=1, aes(size=baseMean), width=0.2)+
  scale_size(guide=FALSE)+
  scale_fill_manual(values = palette_colors)+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=5))+
  theme(axis.title.y = element_text(size=10))+
  theme(legend.position="none")+
  geom_hline(yintercept=0)+
  ylab(expression(Response~to~Chiletanum (log[2]-fold~change)))+
  theme(axis.text.x = element_text(face="italic"),
             legend.text = element_text(face="italic"),
             axis.title.x = element_blank())
C
```


#Modern
```{r}
physeq_Modern = subset_samples(physeq_trimmed, Clade == 'M' & Response != "NA" & Soil_inoculation != "Sterile")
physeq_Modern = subset_samples(physeq_Modern, sample_sums(physeq_Modern) != 0)

diagdds_cladeM = geomMeans_DESeq2(physeq_Modern)
sigtab_M = DESeq_table(diagdds_cladeM,physeq_Modern,0.05)
res_M = res_table(diagdds_cladeM,physeq_Modern)

#At Phylum Level: 
physeq_phylaM = tax_glom(physeq_Modern,taxrank=rank_names(physeq_Modern)[2])
physeq_phylaM = subset_taxa(physeq_phylaM, Phylum != "NA")
physeq_phylaM = subset_samples(physeq_phylaM, sample_sums(physeq_phylaM) != 0)

#At Family Level: 
physeq_FamM = tax_glom(physeq_Modern,taxrank=rank_names(physeq_Modern)[5])
physeq_FamM = subset_taxa(physeq_FamM, Phylum != "NA")
physeq_FamM = subset_samples(physeq_FamM, sample_sums(physeq_FamM) != 0)

#DESeq2 function: 
diagdds_cladeM_Phyla = geomMeans_DESeq2(physeq_phylaM)
diagdds_cladeM_Fam = geomMeans_DESeq2(physeq_FamM)

#Making Tables for significant taxa and all taxas:
sigtab_M_Phyla = DESeq_table(diagdds_cladeM_Phyla,physeq_phylaM,0.05)
sigtab_M_Fam = DESeq_table(diagdds_cladeM_Fam,physeq_FamM,0.05)
res_M_Phyla = res_table(diagdds_cladeM_Phyla,physeq_phylaM)
res_M_Fam = res_table(diagdds_cladeM_Fam,physeq_FamM)

```

```{r}

# Ordering factor by most abundant phyla
res_M = subset(res_M, Phylum != "NA" & log2FoldChange != "NA" & padj != "NA")
TaxOrder = res_M %>%
  group_by(Phylum)%>%
  summarize(Total=n()) %>%
  arrange(-Total)
TaxOrder = TaxOrder$Phylum
res_M$Phylum = ordered(res_M$Phylum, levels=TaxOrder)

library(RColorBrewer)
nb.cols <- 36
palette_colors = colorRampPalette(brewer.pal(8,"Set2"))(nb.cols)

M = ggplot(res_M,aes(x=Phylum, y=log2FoldChange, alpha= if_else(padj < 0.05,20,5), fill=Phylum))+
  geom_jitter(shape=21, stroke=1, aes(size=baseMean), width=0.2)+
  scale_size(guide=FALSE)+
  scale_fill_manual(values = palette_colors)+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=5))+
  theme(axis.title.y = element_text(size=10))+
  theme(legend.position="none")+
  geom_hline(yintercept=0)+
  ylab(expression(Response~to~Modern(log[2]-fold~change)))+
  theme(axis.text.x = element_text(face="italic"),
             legend.text = element_text(face="italic"),
             axis.title.x = element_blank())
M
```


**USE IN PAPER CH2 SI Figure 2**
```{r}
# Ordering factor by most abundant phyla
res_M_Phyla = subset(res_M_Phyla, Phylum != "NA" & log2FoldChange != "NA" & padj != "NA")
TaxOrder = res_M_Phyla %>%
  group_by(Phylum)%>%
  summarize(Total=n()) %>%
  arrange(-Total)
TaxOrder = TaxOrder$Phylum
res_M_Phyla$Phylum = ordered(res_M_Phyla$Phylum, levels=TaxOrder)

library(RColorBrewer)
nb.cols <- 36
palette_colors = colorRampPalette(brewer.pal(8,"Set2"))(nb.cols)

M = ggplot(res_M_Phyla,aes(x=Phylum, y=log2FoldChange, alpha= if_else(pvalue < 0.05,20,5), fill=Phylum))+
  geom_jitter(shape=21, stroke=1, aes(size=baseMean), width=0.2)+
  scale_size(guide=FALSE)+
  scale_fill_manual(values = palette_colors)+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=5))+
  theme(axis.title.y = element_text(size=10))+
  theme(legend.position="none")+
  geom_hline(yintercept=0)+
  ylab(expression(Response~to~Modern(log[2]-fold~change)))+
  theme(axis.text.x = element_text(face="italic"),
             legend.text = element_text(face="italic"),
             axis.title.x = element_blank())
M
```


#Venn Diagram:

##ASVs differentially present VENN (Use in Paper Ch2)
```{r}

#installation guide here: https://github.com/js229/Vennerable
library(Vennerable)
#generate a list of significant ASVs positively enriched:
sigtab_A_P = subset(sigtab_A, log2FoldChange > 0) # 26/37  are positively enriched
sigtab_C_P = subset(sigtab_C, log2FoldChange > 0) # 25/27 are positively enriched
sigtab_M_P = subset(sigtab_M, log2FoldChange > 0) # 118/141 are positively enriched

my_list <- list(sigtab_A_P$otu_id,sigtab_C_P$otu_id,sigtab_M_P$otu_id)
names(my_list) = c("Andigenum", "Chiletanum", "Modern")
Vstem <- Venn(my_list)
Venn_ASV_P = plot(Vstem, doWeights = FALSE)
Venn_ASV_P

```
ASVs positively enriched by Modern, Chiletanum, and Andigenum. 96 ASVs are unique to Modern, 9 for Chiletanum, and 11 for Andigenum. While 7 ASVs are found to be significantly enriched in all three clades. 



Names of Positively enriched "core" where P < 0.05:

```{r}
Vstem@IntersectionSets[["111"]]
```

ASVs to look up: "OTU00205","OTU02013","OTU02463","OTU05272","OTU05982","OTU06044" "OTU08642","OTU08753","OTU09000","OTU09135","OTU09342","OTU09473","OTU09574","OTU09683","OTU10161","OTU10203"


```{r}
#generate a list of significant ASVs negatively enriched:
sigtab_A_N = subset(sigtab_A, log2FoldChange < 0) # 26/37  are positively enriched
sigtab_C_N = subset(sigtab_C, log2FoldChange < 0) # 25/27 are positively enriched
sigtab_M_N = subset(sigtab_M, log2FoldChange < 0) # 118/141 are positively enriched

my_list <- list(sigtab_A_N$otu_id,sigtab_C_N$otu_id,sigtab_M_N$otu_id)
names(my_list) = c("Andigenum", "Chiletanum", "Modern")
Vstem <- Venn(my_list)
Venn_ASV_N = plot(Vstem, doWeights = FALSE)
Venn_ASV_N
```


Let's look more broadly at the family level:

**Use in Paper Ch2**
```{r}
my_list <- list(sigtab_A_N$Family,sigtab_C_N$Family,sigtab_M_N$Family)
names(my_list) = c("Andigenum", "Chiletanum", "Modern")
Vstem <- Venn(my_list)
Venn_ASV_N = plot(Vstem, doWeights = FALSE)
Venn_ASV_N
```
interesting, there's two 2 families unique to Andigenum and Chiletanum. Little overlap in family between modern vs Chiletanum, Andigenum vs Chiletanum amongst statistically significant ASVs grouped by family...in terms of suppression.


```{r}
#generate a list of significant ASVs negatively enriched:
sigtab_A_N = subset(sigtab_A_Phyla, log2FoldChange < 0) # 26/37  are positively enriched
sigtab_C_N = subset(sigtab_C_Phyla, log2FoldChange < 0) # 25/27 are positively enriched
sigtab_M_N = subset(sigtab_M_Phyla, log2FoldChange < 0) # 118/141 are positively enriched

my_list <- list(sigtab_A_N$Phylum,sigtab_C_N$Phylum,sigtab_M_N$Phylum)
names(my_list) = c("Andigenum", "Chiletanum", "Modern")
Vstem <- Venn(my_list)
Venn_ASV_N = plot(Vstem, doWeights = FALSE)
Venn_ASV_N
```
Modern are unique in families of microbes being enriched that is significant (20 vs 1). 8 Families of microbes that are commonly enriched in all three clades. 


```{r}
Vstem@IntersectionSets

```





```{r}
#installation guide here: https://github.com/js229/Vennerable
#generate a list of significant ASVs negatively enriched:
sigtab_A_P = subset(sigtab_A_Phyla, log2FoldChange > 0)
sigtab_C_P = subset(sigtab_C_Phyla, log2FoldChange > 0) 
sigtab_M_P = subset(sigtab_M_Phyla, log2FoldChange > 0) 

my_list <- list(sigtab_A_P$Phylum,sigtab_C_P$Phylum,sigtab_M_P$Phylum)
names(my_list) = c("Andigenum", "Chiletanum", "Modern")
Vstem <- Venn(my_list)
Venn_ASV_P = plot(Vstem, doWeights = FALSE)
Venn_ASV_P

#Do the results look any different if we did tax_glom first and then DESeq2...

```
3 Phyla that are unique to moderns, 1 Phyla that is common to all three clades. 


```{r}
Vstem@IntersectionSets
```
moderns seem to differ in having 3 phyla that are uniquely suppressed that are statistically significant. 


***

#Violin graphs of enriched vs not (not for use in paper):


```{r}
library(RColorBrewer)
library(cowplot)

A <- ggplot(sigtab_A, aes(Phylum, log2FoldChange, fill = Phylum)) +
        geom_violin(scale = "width", adjust = 1, trim = TRUE) +
        scale_fill_brewer(palette = "Accent") +
        #scale_y_continuous(expand = c(0, 0), position="right", labels = function(x)
                           #c(rep(x = "", times = length(x)-2), x[length(x) - 1], "")) +
        theme_cowplot(font_size = 12) +
        theme(legend.position = "none", panel.spacing = unit(0, "lines"),
              panel.background = element_rect(fill = NA, color = "black"),
              strip.background = element_blank(),
              strip.text = element_text(face = "bold"),
              strip.text.y.left = element_text(angle = 0)) +
        ggtitle("Phyla on x-axis") + xlab("Phyla") + ylab("Expression Level")
A

```


```{r}
C <- ggplot(sigtab_C, aes(Phylum, log2FoldChange, fill = Phylum)) +
        geom_violin(scale = "width", adjust = 1, trim = TRUE) +
        scale_fill_brewer(palette = "Accent") +
        #scale_y_continuous(expand = c(0, 0), position="right", labels = function(x)
                           #c(rep(x = "", times = length(x)-2), x[length(x) - 1], "")) +
        theme_cowplot(font_size = 12) +
        theme(legend.position = "none", panel.spacing = unit(0, "lines"),
              panel.background = element_rect(fill = NA, color = "black"),
              strip.background = element_blank(),
              strip.text = element_text(face = "bold"),
              strip.text.y.left = element_text(angle = 0)) +
        ggtitle("Phyla on x-axis") + xlab("Phyla") + ylab("Expression Level")
C
```


***

Graphing overall results: 
```{r}
#Overall results:
x = tapply(res_M$log2FoldChange, res_M$Phylum, function(x) max(x))
x = sort(x, TRUE)
res_M$Phylum= factor(as.character(res_M$Phylum), levels=names(x))
# Genus order
x = tapply(res_M$log2FoldChange, res_M$Genus, function(x) max(x))
x = sort(x, TRUE)
res_M$Genus = factor(as.character(res_M$Genus), levels=names(x))
res_M = res_M

Plot_M = ggplot(sigtab, aes(x=Phylum, y=log2FoldChange, color=Phylum)) + geom_point(size=6) + 
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5))
Plot_M
```



##Scatter plot of general community trends across all phyla:

**Modern**
Focusing on Proteobacteria, Firmicutes, Bacteriodetes, and Actinabacteria:

###Proteobacteria
```{r}
TaxOrder = res_M %>%
    group_by(Phylum) %>%
    group_by(Family)
TaxOrder = TaxOrder$Phylum


library(RColorBrewer)
nb.cols <- 100
palette_colors = colorRampPalette(brewer.pal(8,"Set2"))(nb.cols)

res_M_Proteobacteria = subset(res_M, Phylum == "Proteobacteria")
p = ggplot(res_M_Proteobacteria,aes(x=Family, y=log2FoldChange, alpha= if_else(pvalue < 0.05,20,5), fill=Family))+
  geom_jitter(shape=21, stroke=1, aes(size= 'baseMean'), width=0.2)+
  scale_fill_manual(values = palette_colors)+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=5))+
  #facet_wrap(~Phylum, scales="free_x")+
  theme(legend.position="none")+
  geom_hline(yintercept=0)+
  ylab("Response to Modern Clade (log2-fold change)")
p
```

###Bacteroidota
```{r}
TaxOrder = res_M %>%
    group_by(Phylum) %>%
    group_by(Family)
TaxOrder = TaxOrder$Phylum

nb.cols <- 100
palette_colors = colorRampPalette(brewer.pal(8,"Set2"))(nb.cols)

res_M_Bacteriodetes = subset(res_M, Phylum == "Bacteroidota")
b = ggplot(res_M_Bacteriodetes,aes(x=Family, y=log2FoldChange, alpha= if_else(pvalue < 0.05,20,5), fill=Family))+
  geom_jitter(shape=21, stroke=1, aes(size= 'baseMean'), width=0.2)+
  scale_fill_manual(values = palette_colors)+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=5))+
  #facet_wrap(~Phylum, scales="free_x")+
  theme(legend.position="none")+
  ggtitle("Modern")+
  geom_hline(yintercept=0)+
  ylab("Log2-fold change)")
b
```


###Actinobacteria
```{r}
res_M_Actinobacteriota = subset(res_M, Phylum == "Actinobacteriota")
a = ggplot(res_M_Actinobacteriota,aes(x=Family, y=log2FoldChange, alpha= if_else(pvalue < 0.05,20,5), fill=Family))+
  geom_jitter(shape=21, stroke=1, aes(size= 'baseMean'), width=0.2)+
  scale_fill_manual(values = palette_colors)+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=5))+
  #facet_wrap(~Phylum, scales="free_x")+
  theme(legend.position="none")+
  geom_hline(yintercept=0)+
  ylab("Response to Modern Clade (log2-fold change)")
a
```

###Firmacutes
```{r}
res_M_Firmicutes = subset(res_M, Phylum == "Firmicutes")
f = ggplot(res_M_Firmicutes,aes(x=Family, y=log2FoldChange, alpha= if_else(pvalue < 0.05,20,5), fill=Family))+
  geom_jitter(shape=21, stroke=1, aes(size= 'baseMean'), width=0.2)+
  scale_fill_manual(values = palette_colors)+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=5))+
  #facet_wrap(~Phylum, scales="free_x")+
  theme(legend.position="none")+
  geom_hline(yintercept=0)+
  ylab("Response to Modern Clade (log2-fold change)")
f
```

```{r}
p+a+b+f
```

##Chiletanum:

###Proteobacteria
```{r}
TaxOrder = res_C %>%
    group_by(Phylum) %>%
    group_by(Family)
TaxOrder = TaxOrder$Phylum

library(RColorBrewer)
nb.cols <- 100
palette_colors = colorRampPalette(brewer.pal(8,"Set2"))(nb.cols)

res_C_Proteobacteria = subset(res_C, Phylum == "Proteobacteria")
p = ggplot(res_C_Proteobacteria,aes(x=Family, y=log2FoldChange, alpha= if_else(pvalue < 0.05,20,5), fill=Family))+
  geom_jitter(shape=21, stroke=1, aes(size= 'baseMean'), width=0.2)+
  scale_fill_manual(values = palette_colors)+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=5))+
  #facet_wrap(~Phylum, scales="free_x")+
  theme(legend.position="none")+
  geom_hline(yintercept=0)+
  ggtitle("Chiletanum")+
  ylab("Log2-fold change")
p
```

###Bacteroidota
```{r}
TaxOrder = res_M %>%
    group_by(Phylum) %>%
    group_by(Family)
TaxOrder = TaxOrder$Phylum

nb.cols <- 100
palette_colors = colorRampPalette(brewer.pal(8,"Set2"))(nb.cols)

res_C_Bacteriodetes = subset(res_C, Phylum == "Bacteroidota")
b = ggplot(res_C_Bacteriodetes,aes(x=Family, y=log2FoldChange, alpha= if_else(pvalue < 0.05,20,5), fill=Family))+
  geom_jitter(shape=21, stroke=1, aes(size= 'baseMean'), width=0.2)+
  scale_fill_manual(values = palette_colors)+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=5))+
  #facet_wrap(~Phylum, scales="free_x")+
  theme(legend.position="none")+
  ggtitle("Chiletanum")+
  geom_hline(yintercept=0)+
  ylab("Log2-fold change")
b
```


###Actinobacteria
```{r}
res_C_Actinobacteriota = subset(res_C, Phylum == "Actinobacteriota")
a = ggplot(res_C_Actinobacteriota,aes(x=Family, y=log2FoldChange, alpha= if_else(pvalue < 0.05,20,5), fill=Family))+
  geom_jitter(shape=21, stroke=1, aes(size= 'baseMean'), width=0.2)+
  scale_fill_manual(values = palette_colors)+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=5))+
  #facet_wrap(~Phylum, scales="free_x")+
  theme(legend.position="none")+
  geom_hline(yintercept=0)+
  ggtitle("Chiletanum")+
  ylab("Log2-fold change")
a
```

###Firmacutes
```{r}
res_C_Firmicutes = subset(res_C, Phylum == "Firmicutes")
f = ggplot(res_C_Firmicutes,aes(x=Family, y=log2FoldChange, alpha= if_else(pvalue < 0.05,20,5), fill=Family))+
  geom_jitter(shape=21, stroke=1, aes(size= 'baseMean'), width=0.2)+
  scale_fill_manual(values = palette_colors)+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=5))+
  #facet_wrap(~Phylum, scales="free_x")+
  theme(legend.position="none")+
  geom_hline(yintercept=0)+
  ggtitle("Chiletanum")+
  ylab("Log2-fold change")
f
```



##Andigenum:

```{r}
res_A = res_table(diagdds_cladeA,physeq_Andigenum)
```


```{r}
options(repr.plot.width=14, repr.plot.height=28)

nTax = 200

# Ordering factor by most abundant phyla
TaxOrder = res_A %>%
    group_by(Phylum)
TaxOrder = TaxOrder$Phylum

library(RColorBrewer)
nb.cols <- 36
palette_colors = colorRampPalette(brewer.pal(8,"Set2"))(nb.cols)

A = ggplot(res_A,aes(x=Phylum, y=log2FoldChange, alpha= if_else(pvalue < 0.05,20,5), fill=Phylum))+
  geom_jitter(shape=21, stroke=1, aes(size= 'baseMean'), width=0.2)+
  scale_fill_manual(values = palette_colors)+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=5))+
  #facet_wrap(~Phylum, scales="free_x")+
  theme(legend.position="none")+
  geom_hline(yintercept=0)+
  ylab("Response to Andigenum Clade (log2-fold change)")
A
```


```{r}
A+C+p
```

Let's examine taxa distribution at finer scale (Family):

Here we just based results on non-aggregated ASV:
```{r}
A = ggplot(res_A,aes(x=Family, y=log2FoldChange, alpha= if_else(pvalue < 0.05,20,5), fill=Phylum))+
  geom_jitter(shape=21, stroke=1, aes(size= 'baseMean'), width=0.2)+
  scale_fill_manual(values = palette_colors)+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=5))+
  #facet_wrap(~Phylum, scales="free_x")+
  theme(legend.position="none")+
  geom_hline(yintercept=0)+
  ylab("Response to Andigenum Clade (log2-fold change)")
A


```
Really hard to see...

Here we just based results on aggregated ASV using taxa_glom function in phyloseq:

```{r}
A1 = ggplot(res_A_Fam,aes(x=Family, y=log2FoldChange, alpha= if_else(pvalue < 0.05,20,5), fill=Phylum))+
  geom_jitter(shape=21, stroke=1, aes(size= 'baseMean'), width=0.2)+
  scale_fill_manual(values = palette_colors)+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=5))+
  #facet_wrap(~Phylum, scales="free_x")+
  theme(legend.position="none")+
  geom_hline(yintercept=0)+
  ylab("Response to Andigenum Clade (log2-fold change)")
A1
```
We get two very different graphs for scatter plot aggregated vs non-aggregated reads. Why is that is the case?? and which one should we follow?? I guess its different because each dot in non-aggregated is an ASV while each dot in aggregated is average of the Log2FoldChange of all the ASVs hence why it looks "cleaner". 


```{r}
res_A_Proteobacteria = subset(res_A, Phylum == "Proteobacteria")
p = ggplot(res_A_Proteobacteria,aes(x=Family, y=log2FoldChange, alpha= if_else(pvalue < 0.05,20,5), fill=Family))+
  geom_jitter(shape=21, stroke=1, aes(size= 'baseMean'), width=0.2)+
  scale_fill_manual(values = palette_colors)+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=5))+
  #facet_wrap(~Phylum, scales="free_x")+
  theme(legend.position="none")+
  geom_hline(yintercept=0)+
  ggtitle("Andigenum")+
  ylab("Log2-fold change")
p
```


###Bacteroidota
```{r}
nb.cols <- 100
palette_colors = colorRampPalette(brewer.pal(8,"Set2"))(nb.cols)

res_A_Bacteriodetes = subset(res_A, Phylum == "Bacteroidota")
b = ggplot(res_A_Bacteriodetes,aes(x=Family, y=log2FoldChange, alpha= if_else(pvalue < 0.05,20,5), fill=Family))+
  geom_jitter(shape=21, stroke=1, aes(size= 'baseMean'), width=0.2)+
  scale_fill_manual(values = palette_colors)+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=5))+
  #facet_wrap(~Phylum, scales="free_x")+
  theme(legend.position="none")+
  ggtitle("Andigenum")+
  geom_hline(yintercept=0)+
  ylab("Log2-fold change")
b
```


###Actinobacteria
```{r}
res_A_Actinobacteriota = subset(res_A, Phylum == "Actinobacteriota")
a = ggplot(res_A_Actinobacteriota,aes(x=Family, y=log2FoldChange, alpha= if_else(pvalue < 0.05,20,5), fill=Family))+
  geom_jitter(shape=21, stroke=1, aes(size= 'baseMean'), width=0.2)+
  scale_fill_manual(values = palette_colors)+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=5))+
  #facet_wrap(~Phylum, scales="free_x")+
  theme(legend.position="none")+
  geom_hline(yintercept=0)+
  ggtitle("Andigenum")+
  ylab("Log2-fold change")
a
```

###Firmacutes
```{r}
res_A_Firmicutes = subset(res_A, Phylum == "Firmicutes")
f = ggplot(res_A_Firmicutes,aes(x=Family, y=log2FoldChange, alpha= if_else(pvalue < 0.05,20,5), fill=Family))+
  geom_jitter(shape=21, stroke=1, aes(size= 'baseMean'), width=0.2)+
  scale_fill_manual(values = palette_colors)+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=5))+
  #facet_wrap(~Phylum, scales="free_x")+
  theme(legend.position="none")+
  geom_hline(yintercept=0)+
  ggtitle("Andigenum")+
  ylab("Log2-fold change")
f
```




***

Random things below...


```{r}
res_M1 = subset(res_M)

p = ggplot(res_M,aes(x=Phylum, y=log2FoldChange, alpha= if_else(pvalue < 0.05,20,5), fill=Phylum))+
  geom_jitter(shape=21, stroke=1, aes(size= 'baseMean'), width=0.2)+
  scale_fill_manual(values = palette_colors)+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=5))+
  #facet_wrap(~Phylum, scales="free_x")+
  theme(legend.position="none")+
  geom_hline(yintercept=0)+
  ylab("Response to Modern Clade (log2-fold change)")
p
```


```{r}
C = ggplot(res_C,aes(x=Family, y=log2FoldChange, alpha= if_else(pvalue < 0.05,20,5), fill=Phylum))+
  geom_jitter(shape=21, stroke=1, aes(size= 'baseMean'), width=0.2)+
  scale_fill_manual(values = palette_colors)+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=5))+
  #facet_wrap(~Phylum, scales="free_x")+
  theme(legend.position="none")+
  geom_hline(yintercept=0)+
  ylab("Response to Chiletanum Clade (log2-fold change)")
C
```


```{r}
p = ggplot(res_M,aes(x=Family, y=log2FoldChange, alpha= if_else(pvalue < 0.05,20,5), fill=Phylum))+
  geom_jitter(shape=21, stroke=1, aes(size= 'baseMean'), width=0.2)+
  scale_fill_manual(values = palette_colors)+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=5))+
  #facet_wrap(~Phylum, scales="free_x")+
  theme(legend.position="none")+
  geom_hline(yintercept=0)+
  ylab("Response to Modern Clade (log2-fold change)")
p
```

