---
title: "DESeq2_yield2"
output: html_notebook
---

#Aim 2: Just for the moderns selection, what is different +/- selected soils

**Notes/Rick's comments**: Adonis selection effect on composition. Deseq2 on selection and response. Are you getting the same taxa?…

  + No persistent link between yield and MC for Andigenum and Chiletanum. Curious if there strongly interacting taxa best measure Deseq2 response across all iteration. 

  + Moderns and Hancock: selection worked -> what is different P vs N soils, Stacked bar chart to visualize difference. Deseq2 at Phyla level…

Import required Packages:
```{r}
library(phyloseq)
library(tidyverse)
library(vegan)
library(ggplot2)
library(lme4)
library(lmerTest)
library(emmeans)
library(DESeq2)
```

Import Phyloseq object:
```{r}
physeq_trimmed = readRDS(file = "ps_merged.RDS")
```

Functions to be used for analysis:
```{r}
#Function 1:
geomMeans_DESeq2_Res = function(ps_object){
diagdds_clade= phyloseq_to_deseq2(ps_object, ~ Response)

gm_mean = function(x, na.rm=TRUE){
  exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
}

geoMeans = apply(counts(diagdds_clade), 1, gm_mean)
diagdds = estimateSizeFactors(diagdds_clade, geoMeans = geoMeans)

diagdds_clade = DESeq(diagdds, fitType="parametric")
return(diagdds_clade)
}

#Use Parametric because: There is a capping of gene-wise dispersion values (you probably have a number of genes with all 0's and then a very high count for a single sample), and the parametric trend is less affected by this. The local trend is flattening on the left side.


#Function 2:
geomMeans_DESeq2_Select = function(ps_object){
diagdds_clade= phyloseq_to_deseq2(ps_object, ~ Selection.x)

gm_mean = function(x, na.rm=TRUE){
  exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
}

geoMeans = apply(counts(diagdds_clade), 1, gm_mean)
diagdds = estimateSizeFactors(diagdds_clade, geoMeans = geoMeans)

diagdds_clade = DESeq(diagdds, fitType="parametric")
return(diagdds_clade)
}

#Function 3: significant taxa < alpha
DESeq_table = function(DESeq2_ds,ps_object){
res = results(DESeq2_ds, cooksCutoff = FALSE)
alpha = 0.05
sigtab = res[which(res$padj < alpha), ]
sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(ps_object)[rownames(sigtab), ], "matrix"))
return(sigtab)
}

#Function 4: All results

res_table = function(diagdds_clade,ps_object){
  res = results(diagdds_clade, cooksCutoff = FALSE)
  res_A_Phyla = cbind(as(res, "data.frame"), as(tax_table(ps_object)[rownames(res), ], "matrix"))
}

```



```{r}
#For Genus level treatments:

physeq_ModernH = subset_samples(physeq_trimmed, Clade == 'M' & Response != "NA" & Soil_inoculation != "Sterile" & Iteration == 5 & Soil_type.x == "Hancock")
physeq_ModernH <- tax_glom(physeq_ModernH, taxrank = "Genus", NArm = F)
physeq_ModernH = filter_taxa(physeq_ModernH, function(x) sum(x > 0) > (0.1*length(x)), TRUE)

#At Genus level for Selection:
diagdds_SelectM_ASV = geomMeans_DESeq2_Select(physeq_ModernH)
res_SelectM_GenusH = res_table(diagdds_SelectM_ASV,physeq_ModernH)
sigtab_SelectM_GenusH = res_SelectM_GenusH[which(res_SelectM_GenusH$padj < 0.05), ]

#At Genus level for Response:
diagdds_ResM_ASV = geomMeans_DESeq2_Res(physeq_ModernH)
res_ResM_GenusH = res_table(diagdds_ResM_ASV,physeq_ModernH)
sigtab_ResM_GenusH = res_ResM_GenusH[which(res_ResM_GenusH$padj < 0.05), ]


```

Plot Response:
```{r}
# Ordering factor by logfoldchange in numerical order
TaxOrder = res_ResM_GenusH %>%
  group_by(Phylum)%>%
  summarize(Total=n()) %>%
  arrange(-Total)
TaxOrder = TaxOrder$Phylum
res_ResM_GenusH$Phylum = ordered(res_ResM_GenusH$Phylum, levels=TaxOrder)

nb.cols <- 41 #Number of colors to use for graph for RColorBrewer
palette_colors = colorRampPalette(brewer.pal(8,"Set1"))(nb.cols)

# Making the plot
M_ResH_Plot = ggplot(res_ResM_GenusH,aes(x=Phylum, y=log2FoldChange, alpha= if_else(padj < 0.05,20,5), fill=Phylum))+
  geom_jitter(shape=21, stroke=1, aes(size=baseMean), width=0.2)+
  scale_size(guide=FALSE)+
  scale_fill_manual(values = palette_colors)+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=10))+
  theme(axis.title.y = element_text(size=10))+
  theme(legend.position="none")+
  geom_hline(yintercept=0)+
  ylab(expression(Response~to~Plant~Growth(log[2]-fold~change)))+
  theme(axis.text.x = element_text(face="italic"),
             legend.text = element_text(face="italic"),
             axis.title.x = element_blank())
M_ResH_Plot
```


Plot Selection:
```{r}
TaxOrder = res_SelectM_GenusH %>%
  group_by(Phylum)%>%
  summarize(Total=n()) %>%
  arrange(-Total)
TaxOrder = TaxOrder$Phylum
res_SelectM_GenusH$Phylum = ordered(res_SelectM_GenusH$Phylum, levels=TaxOrder)

nb.cols <- 41 #Number of colors to use for graph for RColorBrewer
palette_colors = colorRampPalette(brewer.pal(8,"Set1"))(nb.cols)

# Making the plot
M_SelectH_Plot = ggplot(res_SelectM_GenusH,aes(x=Phylum, y=log2FoldChange, alpha= if_else(padj < 0.05,20,5), fill=Phylum))+
  geom_jitter(shape=21, stroke=1, aes(size=baseMean), width=0.2)+
  scale_size(guide=FALSE)+
  scale_fill_manual(values = palette_colors)+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=10))+
  theme(axis.title.y = element_text(size=10))+
  theme(legend.position="none")+
  geom_hline(yintercept=0)+
  ylab(expression(Response~to~Selection(log[2]-fold~change)))+
  theme(axis.text.x = element_text(face="italic"),
             legend.text = element_text(face="italic"),
             axis.title.x = element_blank())
M_SelectH_Plot
```


```{r}
#installation guide here: https://github.com/js229/Vennerable
library(Vennerable)

restab_ResM_Genus_P = subset(res_ResM_GenusH, log2FoldChange > 0)
restab_SelectM_P = subset(res_SelectM_GenusH, log2FoldChange > 0)

#generate a list of significant ASVs that are were enriched
my_list <- list(restab_ResM_Genus_P$Genus,restab_SelectM_P$Genus)
names(my_list) = c("Response","Selection")
Vstem <- Venn(my_list)
plot(Vstem, doWeights = FALSE)
```


```{r}
Vstem@IntersectionSets[["11"]]
```



~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

#Phyla:

```{r}
#Phyloseq averaging of abundance for different phylogeny levels
physeq_ModernH = subset_samples(physeq_trimmed, Clade == 'M' & Response != "NA" & Soil_inoculation != "Sterile" & Iteration == 5 & Soil_type.x == "Hancock")
physeq_phylaM_H = tax_glom(physeq_ModernH,taxrank=rank_names(physeq_ModernH)[2])
physeq_phylaM_H = filter_taxa(physeq_phylaM_H, function(x) sum(x > 0) > (0.1*length(x)), TRUE)


#DESeq2 function: 
diagdds_ResM_Phyla = geomMeans_DESeq2_Res(physeq_phylaM_H)
diagdds_SelectM_Phyla = geomMeans_DESeq2_Select(physeq_phylaM_H)

#Making Tables for all taxa 
res_ResM_PhylaH = res_table(diagdds_ResM_Phyla,physeq_phylaM_H)
res_ResM_PhylaH = subset(res_ResM_PhylaH, log2FoldChange != "NA")

res_SelectM_PhylaH = res_table(diagdds_SelectM_Phyla,physeq_phylaM_H)
res_SelectM_PhylaH = subset(res_SelectM_PhylaH, log2FoldChange != "NA")

#Making Tables for significant taxa only:
sigtab_SelectM_PhylaH = res_SelectM_PhylaH[which(res_SelectM_PhylaH$padj < 0.05),]
sigtab_ResM_PhylaH = res_ResM_PhylaH[which(res_ResM_PhylaH$padj < 0.05),]

```


```{r}
#installation guide here: https://github.com/js229/Vennerable
library(Vennerable)

restab_ResM_Phyla_P = subset(res_ResM_PhylaH, log2FoldChange > 0)
restab_SelectM_Phyla_P = subset(res_SelectM_PhylaH, log2FoldChange > 0)

#generate a list of significant ASVs that are were enriched
my_list <- list(restab_ResM_Phyla_P$Phylum,restab_SelectM_Phyla_P$Phylum)
names(my_list) = c("Response","Selection")
Vstem <- Venn(my_list)
plot(Vstem, doWeights = FALSE)
```


Visualizing Results:

**Response**
```{r}
TaxOrder = res_ResM_PhylaH %>%
  group_by(Phylum)%>%
  arrange(-baseMean)
TaxOrder = TaxOrder$Phylum
res_ResM_PhylaH$Phylum = ordered(res_ResM_PhylaH$Phylum, levels=TaxOrder)

nb.cols <- 41 #Number of colors to use for graph for RColorBrewer
palette_colors = colorRampPalette(brewer.pal(8,"Set1"))(nb.cols)

# Making the plot
M_ResH_PhylaPlot = ggplot(res_ResM_PhylaH,aes(x=Phylum, y=log2FoldChange, alpha= if_else(padj < 0.05,20,5), fill=Phylum))+
  geom_jitter(shape=21, stroke=1, aes(size=baseMean), width=0.2)+
  scale_size(guide=FALSE)+
  scale_fill_manual(values = palette_colors)+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=10))+
  theme(axis.title.y = element_text(size=10))+
  theme(legend.position="none")+
  geom_hline(yintercept=0)+
  ylab(expression(Response~to~Plant~Growth(log[2]-fold~change)))+
  theme(axis.text.x = element_text(face="italic"),
             legend.text = element_text(face="italic"),
             axis.title.x = element_blank())
M_ResH_PhylaPlot
```
Moderns see broader and larger number of ASVs differentially abundant with Plant response. 

```{r}
#Phyloseq averaging of abundance for different phylogeny levels
physeq_ModernW = subset_samples(physeq_trimmed, Clade == 'M' & Response != "NA" & Soil_inoculation != "Sterile" & Iteration == 5 & Soil_type.x == "Winterfell")
physeq_phylaM_W = tax_glom(physeq_ModernH,taxrank=rank_names(physeq_ModernH)[2])
physeq_phylaM_W = filter_taxa(physeq_phylaM_H, function(x) sum(x > 0) > (0.1*length(x)), TRUE)

#DESeq2 function: 
diagdds_ResM_Phyla = geomMeans_DESeq2_Res(physeq_phylaM_W)
diagdds_SelectM_Phyla = geomMeans_DESeq2_Select(physeq_phylaM_W)

#Making Tables for all taxa 
res_ResM_PhylaW = res_table(diagdds_ResM_Phyla,physeq_phylaM_W)
res_ResM_PhylaW = subset(res_ResM_PhylaH, log2FoldChange != "NA")

res_SelectM_PhylaW = res_table(diagdds_SelectM_Phyla,physeq_phylaM_W)
res_SelectM_PhylaW = subset(res_SelectM_PhylaH, log2FoldChange != "NA")

#Making Tables for significant taxa only:
sigtab_SelectM_PhylaW = res_SelectM_PhylaH[which(res_SelectM_PhylaH$padj < 0.05),]
sigtab_ResM_PhylaW = res_ResM_PhylaH[which(res_ResM_PhylaH$padj < 0.05),]

```


```{r}
TaxOrder = res_ResM_PhylaW %>%
  group_by(Phylum)%>%
  arrange(-baseMean)
TaxOrder = TaxOrder$Phylum
res_ResM_PhylaW$Phylum = ordered(res_ResM_PhylaW$Phylum, levels=TaxOrder)

nb.cols <- 41 #Number of colors to use for graph for RColorBrewer
palette_colors = colorRampPalette(brewer.pal(8,"Set1"))(nb.cols)

# Making the plot
M_ResW_PhylaPlot = ggplot(res_ResM_PhylaW,aes(x=Phylum, y=log2FoldChange, alpha= if_else(padj < 0.05,20,5), fill=Phylum))+
  geom_jitter(shape=21, stroke=1, aes(size=baseMean), width=0.2)+
  scale_size(guide=FALSE)+
  scale_fill_manual(values = palette_colors)+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=10))+
  theme(axis.title.y = element_text(size=10))+
  theme(legend.position="none")+
  geom_hline(yintercept=0)+
  ylab(expression(Response~to~Plant~Growth(log[2]-fold~change)))+
  theme(axis.text.x = element_text(face="italic"),
             legend.text = element_text(face="italic"),
             axis.title.x = element_blank())
M_ResW_PhylaPlot
```



**Selection**
```{r}
res_SelectM_PhylaW = subset(res_SelectM_PhylaW, Phylum != "NA" & log2FoldChange != "NA" & padj != "NA")
TaxOrder = res_SelectM_PhylaW %>%
  arrange(-baseMean)
TaxOrder = TaxOrder$Phylum
res_SelectM_PhylaW$Phylum = ordered(res_SelectM_PhylaW$Phylum, levels=TaxOrder)

nb.cols <- 41 #Number of colors to use for graph for RColorBrewer
palette_colors = colorRampPalette(brewer.pal(8,"Set2"))(nb.cols)

# Making the plot
M_SelectW_PhylaPlot = ggplot(res_SelectM_PhylaW,aes(x=Phylum, y=log2FoldChange, alpha= if_else(padj < 0.05,20,5), fill=Phylum))+
  geom_jitter(shape=21, stroke=1, aes(size=baseMean), width=0.2)+
  scale_size(guide=FALSE)+
  scale_fill_manual(values = palette_colors)+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=10))+
  theme(axis.title.y = element_text(size=10))+
  theme(legend.position="none")+
  geom_hline(yintercept=0)+
  ylab(expression(Response~to~Selection(log[2]-fold~change)))+
  theme(axis.text.x = element_text(face="italic"),
             legend.text = element_text(face="italic"),
             axis.title.x = element_blank())
M_SelectW_PhylaPlot
```


```{r}
#installation guide here: https://github.com/js229/Vennerable
library(Vennerable)

sigtab_ResM_ASV_P = subset(res_ResM_PhylaW, log2FoldChange > 0)
sigtab_M_Select_P = subset(res_SelectM_PhylaW, log2FoldChange > 0)

#generate a list of significant ASVs that are were enriched
my_list <- list(sigtab_ResM_ASV_P$Phylum,sigtab_M_Select_P$Phylum)
names(my_list) = c("Response","Selected")
Vstem <- Venn(my_list)
plot(Vstem, doWeights = FALSE)
```

```{r}
Vstem@IntersectionSets
```




```{r}
#installation guide here: https://github.com/js229/Vennerable
library(Vennerable)

sigtab_ResM_ASV_N = subset(sigtab_ResM_ASV, log2FoldChange < 0)
sigtab_M_Select_N = subset(sigtab_SelectM_ASV, log2FoldChange < 0)

#generate a list of significant ASVs that are were enriched
my_list <- list(sigtab_ResM_ASV_N$Phyla,sigtab_M_Select_N$Family)
names(my_list) = c("Response","Selected")
Vstem <- Venn(my_list)
plot(Vstem, doWeights = FALSE)
```

```{r}
Vstem@IntersectionSets
```

***



```{r}
#Phyloseq averaging of abundance for different phylogeny levels
physeq_ModernW = subset_samples(physeq_trimmed, Clade == 'M' & Response != "NA" & Soil_inoculation != "Sterile" & Iteration == 5 & Soil_type.x == "Winterfell")
physeq_phylaM_W = tax_glom(physeq_ModernW,taxrank=rank_names(physeq_ModernW)[2])
physeq_phylaM_W = filter_taxa(physeq_phylaM_W, function(x) sum(x > 0) > (0.1*length(x)), TRUE)


#DESeq2 function: 
diagdds_ResM_Phyla = geomMeans_DESeq2_Res(physeq_phylaM_W)
diagdds_SelectM_Phyla = geomMeans_DESeq2_Select(physeq_phylaM_W)

#Making Tables for all taxa 
res_ResM_PhylaW = res_table(diagdds_ResM_Phyla,physeq_phylaM_W)
res_ResM_PhylaW = subset(res_ResM_PhylaW, log2FoldChange != "NA")

res_SelectM_PhylaW = res_table(diagdds_SelectM_Phyla,physeq_phylaM_W)
res_SelectM_PhylaW = subset(res_SelectM_PhylaW, log2FoldChange != "NA")

#Making Tables for significant taxa only:
sigtab_SelectM_PhylaW = res_SelectM_PhylaW[which(res_SelectM_PhylaW$padj < 0.05),]
sigtab_ResM_PhylaW = res_ResM_PhylaW[which(res_ResM_PhylaW$padj < 0.05),]


```

##Positive/Negative Selection In Winterfell
```{r}
library(RColorBrewer)

# Ordering factor by logfoldchange in numerical order
res_SelectM_PhylaW = subset(res_SelectM_PhylaW, Order != "NA" & log2FoldChange != "NA" & padj != "NA")
TaxOrder = res_SelectM_PhylaW %>%
  group_by(Phylum)%>%
  summarize(Total=n()) %>%
  arrange(-Total)
TaxOrder = TaxOrder$Phylum
res_SelectM_PhylaW$Phylum = ordered(res_SelectM_PhylaW$Phylum, levels=TaxOrder)

nb.cols <- 41 #Number of colors to use for graph for RColorBrewer
palette_colors = colorRampPalette(brewer.pal(8,"Set2"))(nb.cols)

# Making the plot
W_SelectM_Plot = ggplot(res_SelectM_PhylaW,aes(x=Phylum, y=log2FoldChange, alpha= if_else(padj < 0.05,20,5), fill=Phylum))+
  geom_jitter(shape=21, stroke=1, aes(size=baseMean), width=0.2)+
  scale_size(guide=FALSE)+
  scale_fill_manual(values = palette_colors)+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=10))+
  theme(axis.title.y = element_text(size=10))+
  theme(legend.position="none")+
  geom_hline(yintercept=0)+
  ylab(expression(Response~to~Plant~Growth(log[2]-fold~change)))+
  theme(axis.text.x = element_text(face="italic"),
             legend.text = element_text(face="italic"),
             axis.title.x = element_blank())
W_SelectM_Plot
```
##Response and Negative Selection

```{r}
# Ordering factor by logfoldchange in numerical order
res_ResH_N = subset(res_ResH_N, Order != "NA" & log2FoldChange != "NA" & padj != "NA")
TaxOrder = res_ResH_N %>%
  group_by(Order)%>%
  summarize(Total=n()) %>%
  arrange(-Total)
TaxOrder = TaxOrder$Order[1:69]
res_ResH_N$Order = ordered(res_ResH_N$Order, levels=TaxOrder)

nb.cols <- 69 #Number of colors to use for graph for RColorBrewer
palette_colors = colorRampPalette(brewer.pal(8,"Set2"))(nb.cols)

# Making the plot
H_ResN_Plot = ggplot(res_ResH_N,aes(x=Order, y=log2FoldChange, alpha= if_else(padj < 0.05,20,5), fill=Order))+
  geom_jitter(shape=21, stroke=1, aes(size=baseMean), width=0.2)+
  scale_size(guide=FALSE)+
  scale_fill_manual(values = palette_colors)+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=10))+
  theme(axis.title.y = element_text(size=10))+
  theme(legend.position="none")+
  geom_hline(yintercept=0)+
  ylab(expression(Response~to~Plant~Growth(log[2]-fold~change)))+
  theme(axis.text.x = element_text(face="italic"),
             legend.text = element_text(face="italic"),
             axis.title.x = element_blank())
H_ResN_Plot
```

```{r}
library(Vennerable)

sigtab_ResH_P1 = subset(sigtab_ResH_P, log2FoldChange > 0)
sigtab_ResH_N1 = subset(sigtab_ResH_N, log2FoldChange > 0)


#generate a list of significant ASVs that are were enriched
my_list <- list(sigtab_ResH_P1$Family,sigtab_ResH_N$Family)
names(my_list) = c("Positive","Negative")
Vstem <- Venn(my_list)
plot(Vstem, doWeights = FALSE)
```

```{r}
Vstem@IntersectionSets

```
1] "Caulobacteraceae"   "Longimicrobiaceae"  "Sphingomonadaceae" 
[4] "Micromonosporaceae"

are our microbial families that we can say may have had a positive impact on plant growth and were selected for from our greenhouse experiment.


```{r}
sigtab_ResH_P1 = subset(sigtab_ResH_P, log2FoldChange < 0)
sigtab_ResH_N1 = subset(sigtab_ResH_N, log2FoldChange < 0)


#generate a list of significant ASVs that are were enriched
my_list <- list(sigtab_ResH_P1$Family,sigtab_ResH_N$Family)
names(my_list) = c("Positive","Negative")
Vstem <- Venn(my_list)
plot(Vstem, doWeights = FALSE)

```

```{r}
Vstem@IntersectionSets

```
"0319-6G20"         "Sphingomonadaceae" "Devosiaceae"      
"Paenibacillaceae"  "Dojkabacteria"     "Oxalobacteraceae" 
"Comamonadaceae"    "Saccharimonadales" "Nitrosomonadaceae"
"Rhizobiaceae"      "Flavobacteriaceae" "Planococcaceae"   
"Nocardioidaceae"   "Microbacteriaceae" "Xanthomonadaceae" 

Depletion of these taxa due to our Selection led to worse plant outcomes indicated by our plant response phenotype in Hancock soils. 


***
##Selection
```{r}
# Ordering factor by logfoldchange in numerical order
res_H_ASV1 = subset(res_H_ASV1, Order != "NA" & log2FoldChange != "NA" & padj != "NA")
TaxOrder = res_H_ASV1 %>%
  group_by(Order)%>%
  summarize(Total=n()) %>%
  arrange(-Total)
TaxOrder = TaxOrder$Order[1:69]
res_H_ASV1$Order = ordered(res_H_ASV1$Order, levels=TaxOrder)

nb.cols <- 69 #Number of colors to use for graph for RColorBrewer
palette_colors = colorRampPalette(brewer.pal(8,"Set2"))(nb.cols)

# Making the plot
H_Select_ASV = ggplot(res_H_ASV1,aes(x=Order, y=log2FoldChange, alpha= if_else(padj < 0.05,20,5), fill=Order))+
  geom_jitter(shape=21, stroke=1, aes(size=baseMean), width=0.2)+
  scale_size(guide=FALSE)+
  scale_fill_manual(values = palette_colors)+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=10))+
  theme(axis.title.y = element_text(size=10))+
  theme(legend.position="none")+
  geom_hline(yintercept=0)+
  ylab(expression(Response~to~Plant~Growth(log[2]-fold~change)))+
  theme(axis.text.x = element_text(face="italic"),
             legend.text = element_text(face="italic"),
             axis.title.x = element_blank())
H_Select_ASV
```
Selection and Response are linked sooo we see the sample results....

