---
title: "16S analysis: Figure 1A-C"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

**Loading packages for Use**
```{r}
library(dplyr)
library(vegan)
library(tidyr)
library(ggplot2)
library(readr)
library(reshape)
```

**Formatting Data**
```{r}
taxa_table = read_tsv('merged-taxonomy.tsv')
taxa_table2 = separate(taxa_table,taxonomy, sep = ";", into=c("Kingdom","Phylum","Class","Order","Family","Genus","Species"))
ASV_table = read_tsv('16S-merged-table.txt')

#Note: for the ASV_table we modified it to remove the first row containing "Converted from biom file" for better data carving later...

```

Note: on excel we made some changes to the taxa ('taxonomy.tsv') table prior to uploading taxa table onto R

```{r}
taxa_table2$`OTU ID` <- as.character(taxa_table$`#OTUID`) # Must be same type for join to work

ASV_table$`#OTU_ID` <- as.character(ASV_table$`#OTU_ID`) # Must be same type for join to work
otu_data <- right_join(ASV_table, taxa_table2,
                      by = c("#OTU_ID" = "#OTUID")) # identifies cols with shared IDs

otu_data2 <- right_join(taxa_table2,ASV_table,
                      by = c("#OTUID"="#OTU_ID")) # identifies cols with shared IDs
print(otu_data)
```

```{r}
tail(colnames(otu_data2), n = 10)
head(colnames(otu_data2), n = 11)
```

Now we have made a table that has taxonomy information and also our ASV information
If we want to now include our metadata we have couple of options:
1) phyloseq option 
2) taxa 
3) Rick's method using vegan 

For all intensive purposes, I am going to proceed analyzing my data using Vegan and formatting it accordingly, and formatting my data tables for use in vegan

Reformatting data for use for vegan and other downstream analysis:
```{r}
library(reshape2)

#Use dplry here to reformat table for reshaping:
otu_data3 <- select(otu_data2,'#OTUID','Kingdom','Phylum','Class','Order','Family','Genus','Species',H2_1_5:WA2_9_3)

#reshaping table: 
melt_OTU <- melt(otu_data3[1:436], id.vars = colnames(otu_data3[1:8]))
colnames(otu_data3[1:8])

ASVcast = dcast(melt_OTU, Kingdom + Phylum + Class + Order + Family + Genus + Species ~ variable, fun.aggregate = sum)
phygen = paste(ASVcast$Kingdom, ASVcast$Phylum, ASVcast$Class,ASVcast$Order, ASVcast$Family, ASVcast$Genus,ASVcast$Species)
ASVcast2 = cbind(phygen,ASVcast[,8:435])
ASVcast3 = t(ASVcast2)
ASVcast3  = as.data.frame(ASVcast3)

#Notice the header is incorrect V1...V371
names(ASVcast3) = c(phygen)
ASVcast3 = ASVcast3[-1,]

write.csv(ASVcast3, file="ghexpASV_rdp.csv")

```


Reformatting and Merging ASV data with Metadata: 
```{r}
metadata = read.csv('metadata3.csv', header = T)
metadata = unite(metadata,Clade_Soil_Time, c('Clade','Soil_type.x','Iteration'), sep = '_', remove = F)
ASVdata= read.csv('ghexpASV_rdp.csv', header = T)

Gh18_data = left_join(metadata,ASVdata, by=c("Submission_ID"="X"))

#Now we need to clean up our table for analysis again:

#Remove rows with NA values for ASV data
Gh18_data = drop_na(Gh18_data,d__Bacteria..p__Actinobacteriota..c__Actinobacteria..o__Streptomycetales..f__Streptomycetaceae..g__Streptomyces.NA)

write.csv(Gh18_data, file="Gh18_ASV_data.csv")

#remove columns with the containing the name 'mitochondria' and 'chloroplast'
Gh18_data = Gh18_data[,!grepl("*mitochondria",names(Gh18_data))]
Gh18_data = Gh18_data[,!grepl("*chloroplast",names(Gh18_data))]
write.csv(Gh18_data, file="Gh18_ASV_data.csv")
#Gh18_data = read.csv("Gh18_ASV_data.csv", header =T)

```

Now we are done reformatting and tidying up our data. We will now begin to analyze our data in R using Vegan: 

```{r}
library(vegan)
rarecurve(Gh18_data[32:1695], col = "blue")
```
Based on our rarefaction plot sample size of at least 2000(-ish) sequences per sample to have enough coverage of the microbial community. 


created new column that gives us the sum of sequences: 
```{r}
Gh18_data1 = cbind(Gh18_data,total=rowSums(Gh18_data[32:1695]))

#Quality control step:
Gh18_data2 = subset(Gh18_data1,total>2000)
```

#Beta Diversity Communith Analysis: 

Now we will relativize our data:

```{r}
Gh18_data3 = decostand(Gh18_data2[32:1695], method = "total")
Gh18_data3= bind_cols(Gh18_data2[1:31],Gh18_data3)
#write.csv(Gh18_data3, file="Gh18_ASV_reldata.csv")


set.seed(1001)
ghsoilmds <- metaMDS(Gh18_data3[32:1695], distance = "bray", k = 2)
#plot(ghsoilmds, display = "species", type= "n")
#points(ghsoilmds,display = "sites", col = as.numeric(Gh18_data3$Clade)) #ordiellipse(ghsoilmds, Gh18_data3$Clade, label = F, draw = "polygon", col = c("Black","Blue","Red", "Purple"), alpha = 70)
```
We have very low stress for our data which is interesting k=0.0005436404. Pretty good...but also NMDS may not be appropriate for our data...let's try ordinating with PCoA method...

```{r}
dist <- vegdist(Gh18_data3[32:1695],method = "bray")
library(ape)
PCOA <- pcoa(dist)
eigen_plot <- barplot(PCOA$values$Relative_eig[1:10])

```
Our eignenvalue plot shows that the first two PCoA ordinates are sufficient for explaining 30% of the variance....hmmmm we'll see how this goes.

Axis 1 is 14.3%, and Axis2 is 8.5% = 22.7% of variance is captured which isn't great but also best we can do....

```{r}
PCOA <- pcoa(dist, correction = "cailliez")
biplot.pcoa(PCOA)
PCoA_GH <- cbind(Gh18_data3[1:31],PCOA[["vectors"]])

Gh18_PCoA = ggplot(data = PCoA_GH, aes(Axis.1,Axis.2))+
    #geom_point(aes(color=Clade,shape=Selection.x))+
    stat_ellipse(aes(fill = Clade_Soil_Time), geom="polygon",level=0.95,alpha=0.2) +
    theme_bw()
Gh18_PCoA
```

#Let's try to seperate these now by Clade and replot:

##Andigenum
```{r}
library(RColorBrewer)
nb.cols <- 10 #Number of colors to use for graph for RColorBrewer
gradient = 

palette_colors = colorRampPalette(brewer.pal(8,"Set1"))(nb.cols)
PCoA_GH_A = subset(PCoA_GH, Clade == "A" & Soil_type.x != "Sterile" & Iteration != 2 & Iteration != 4)

Gh18_PCoA = ggplot(data = PCoA_GH_A, aes(Axis.1,Axis.2))+
        geom_point(aes(color=Iteration,shape=Selection.x))+
    stat_ellipse(aes(fill = Clade_Soil_Time, group = Clade_Soil_Time), geom="polygon",level=0.95,alpha= 0.25) +
    theme_bw()
Gh18_PCoA

```

```{r}
set.seed(3001)
Gh18_data_A = subset(Gh18_data3, Clade == "A" & Soil_type.x != "Sterile" & Selection.x != "none")
fit_A = adonis(Gh18_data_A[32:1695] ~ Soil_type.x * Selection.x * Iteration, data = Gh18_data_A, permutations = 999)
fit_A
```


##Chiletanum
```{r}
library(RColorBrewer)
nb.cols <- 10 #Number of colors to use for graph for RColorBrewer
gradient = 

palette_colors = colorRampPalette(brewer.pal(8,"Set1"))(nb.cols)
PCoA_GH_C = subset(PCoA_GH, Clade == "C" & Soil_type.x != "Sterile" & Iteration != 2 & Iteration != 4)

Gh18_PCoA = ggplot(data = PCoA_GH_C, aes(Axis.1,Axis.2))+
        geom_point(aes(color=Iteration,shape=Selection.x))+
    stat_ellipse(aes(fill = Clade_Soil_Time, group = Clade_Soil_Time), geom="polygon",level=0.95,alpha= 0.25) +
    theme_bw()
Gh18_PCoA
```

```{r}
set.seed(3002)
Gh18_data_C = subset(Gh18_data3, Clade == "C" & Soil_type.x != "Sterile" & Selection.x != "none")
fit_C = adonis(Gh18_data_C[32:1695] ~ Soil_type.x * Selection.x * Iteration, data = Gh18_data_C, permutations = 999)
fit_C
```
See Strong soil effect on the microbial community. (Treatment)
See Strong Iteration effect on the microbial community. (Stocasticity)

No interaction effect. So probably means that Soil types fixed in direction and stayed in that general trajectory. 

##Moderns
```{r}
library(RColorBrewer)
nb.cols <- 10 #Number of colors to use for graph for RColorBrewer
gradient = 

palette_colors = colorRampPalette(brewer.pal(8,"Set1"))(nb.cols)
PCoA_GH_M = subset(PCoA_GH, Clade == "M" & Soil_type.x != "Sterile" & Iteration != 2 & Iteration != 4)

Gh18_PCoA = ggplot(data = PCoA_GH_M, aes(Axis.1,Axis.2))+
    geom_point(aes(color=Iteration,shape=Selection.x))+
    stat_ellipse(aes(fill = Clade_Soil_Time, group = Clade_Soil_Time), geom="polygon",level=0.95,alpha= 0.25) +
    theme_bw()
Gh18_PCoA
```

```{r}
set.seed(3003)
Gh18_data_M = subset(Gh18_data3, Clade == "M" & Soil_type.x != "Sterile")
fit_M = adonis(Gh18_data_M[32:1695] ~ Soil_type.x * Selection.x * Iteration, data = Gh18_data_M, permutations = 999)
fit_M
```


##Beta diversity at iteration 5
```{r}
set.seed(1002)
Gh18_data_I5 = subset(Gh18_data3, Iteration == 5 & Soil_type.x != "Sterile")

#ghsoilmds <- metaMDS(Gh18_data_I5[31:1694], distance = "bray", k = 2)
#plot(ghsoilmds, display = "species", type= "n")
#points(ghsoilmds,display = "sites", col = as.numeric(Gh18_data_I5$Clade), bg = ordiellipse(ghsoilmds, Gh18_data_I5$Clade, label = T, draw = "polygon", col = c("Black","Blue","Red", "Purple", alpha = 70))
```


##Graph of Iteration 5
```{r}
Gh18_PCoA_I5 = subset(PCoA_GH, Iteration == 5 & Selection.x != "S")
#Gh18_PCoA_I5 = unite(Gh18_PCoA_I5,Clade,Soil_type.x,col = "Clade_Soil",sep = "_",remove = FALSE)

Gh18_data_I5 = subset(Gh18_data3, Iteration == 5 & Soil_type.x != "Sterile")

ghsoilmds <- metaMDS(Gh18_data_I5[32:1695], distance = "bray", k = 2)
Gh18_MDS_I5 = as.data.frame(ghsoilmds[["points"]])
Gh18_NMDS_I5 = bind_cols(Gh18_PCoA_I5[1:31],Gh18_MDS_I5)
Gh18_Ord_I5 = ggplot(data = Gh18_NMDS_I5, aes(MDS1,MDS2))+
    geom_point(aes(color=Clade,shape=Selection.x))+
    #stat_ellipse(aes(fill = Clade), geom="polygon",level=0.95,alpha=0.2) +
    theme_bw()
Gh18_Ord_I5 


Gh18_Plot_I5 = ggplot(data = Gh18_PCoA_I5 , aes(Axis.1,Axis.2))+
    geom_point(aes(color=Clade,shape=Soil_type.x))+
    #geom_polygon(aes(fill = Clade_Soil_Time), alpha = 0.5) +
    labs(x = "PCoA 1 (13.4%)", y = "PCoA 2 (8.4%)")+
    theme_bw()
Gh18_Plot_I5 
```

###Models for iteration 5:
```{r}
Gh18_data_I5 = subset(Gh18_data3, Iteration == 5 & Soil_type.x != "Sterile")

set.seed(1010)
fit_I5 = adonis(Gh18_data_I5[32:1694] ~ Soil_type.x * Clade * Selection.x, data = Gh18_data_I5, permutations = 999)
fit_I5

set.seed(1011)
Gh18_data_I5_M = subset(Gh18_data_I5, Clade == "M")
fit_I5_M = adonis(Gh18_data_I5_M [32:1694] ~ Soil_type.x * Selection.x, data = subset(Gh18_data_I5_M, Clade == "M"), permutations = 999)
fit_I5_M

set.seed(1012)
Gh18_data_I5_C = subset(Gh18_data_I5, Clade == "C")
fit_I5_C = adonis(Gh18_data_I5_C[32:1694] ~ Soil_type.x * Selection.x, data = Gh18_data_I5_C , permutations = 999)
fit_I5_C

set.seed(1013)
Gh18_data_I5_A = subset(Gh18_data_I5, Clade == "A")
fit_I5_A = adonis(Gh18_data_I5_A[32:1694] ~ Soil_type.x * Selection.x, data = Gh18_data_I5_A, permutations = 999)
fit_I5_A

```
We see a very strong Clade effect! Modern, Chiletanum, and Andigenum are recruiting 3 different kinds of microbial communities. Hence why the p-value = $0.001$

We also see selection playing a role depending on soil type...for Andigenum and Modern.


> Note: Look at stratification concepts with perMANOVA....read/find time series with microbiome work...
>RA: Look at potato phenotype data too...(P,N, vs Sterile); DONE

```{r}
pairwise.adonis2(Gh18_data_I5[32:1695] ~ Clade*Soil_type.x, data = Gh18_data_I5)
```



##Beta diversity at iteration 1
```{r}
set.seed(1003)
Gh18_data_I1 = subset(Gh18_data3, Iteration == 1 & Soil_type.x != "Sterile")
ghsoilmds <- metaMDS(Gh18_data_I1[32:1695], distance = "bray", k = 2)
Gh18_MDS_I1 = as.data.frame(ghsoilmds[["points"]])
Gh18_Ord_I1 = bind_cols(Gh18_data_I1[1:31],Gh18_MDS_I1)
Gh18_PCoA_I1 = subset(PCoA_GH, Iteration == 1 & Selection.x != "S")

#NMDS stuff that we are ignoring:

#ghsoilmds <- metaMDS(Gh18_PCoA_I1 [31:740], distance = "bray", k = 2)
#plot(ghsoilmds, display = "species", type= "n")
#points(ghsoilmds,display = "sites", col = as.numeric(Gh18_data_I1$Clade), bg = Gh18_data_I5$Clade, pch = as.numeric(Gh18_data_I1$Iteration)+ 14)
#ordiellipse(ghsoilmds, Gh18_data_I1$Clade, label = T, draw = "polygon", col = c("Black","Blue","Red", "Purple", alpha = 70))
```

Look at the marginal residuals, want to split dataset out. 

###Plotting with ggplot:
```{r}
#NMDS:
Gh18_MDS_I1 = as.data.frame(ghsoilmds[["points"]])
Gh18_Plot_I1 = bind_cols(Gh18_PCoA_I1[1:31],Gh18_MDS_I1)
Gh18_Ord_I1 = ggplot(data = Gh18_Plot_I1, aes(MDS1,MDS2))+
    geom_point(aes(color=Clade,shape=Selection.x))+
    #stat_ellipse(aes(fill = Clade), geom="polygon",level=0.95,alpha=0.2) +
    theme_bw()
Gh18_Ord_I1 

#PCoA Plot (Use for Figure 1A)

Gh18_Plot_I1 = ggplot(data = Gh18_PCoA_I1 , aes(Axis.1,Axis.2))+
    geom_point(aes(color=Clade,shape=Soil_type.x))+
    #stat_ellipse(aes(fill = Clade), geom="polygon",level=0.95,alpha=0.2) +
    labs(x = "PCoA 1 (13.4%)", y = "PCoA 2 (8.4%)")+
    theme_bw()
Gh18_Plot_I1 
```

##Model
```{r}
fit_I1 = adonis(Gh18_data_I1[32:1694] ~ Soil_type.x * Clade, data = Gh18_data_I1, permutations = 999)
fit_I1

```

## Beta diveristy within iteration 3
```{r}
set.seed(1004)
Gh18_data_I3 = subset(Gh18_data3, Iteration == 3 & Soil_type.x != "Sterile")
ghsoilmds <- metaMDS(Gh18_data_I3[32:1695], distance = "bray", k = 2)
#plot(ghsoilmds, display = "species", type= "n")
#points(ghsoilmds,display = "sites", col = as.numeric(Gh18_data_I3$Clade), bg = Gh18_data_I3$Soil.type...2, pch = as.numeric(Gh18_data_I3$Selection)+ 14)
#ordiellipse(ghsoilmds, Gh18_data_I3$Clade, label = T, draw = "polygon", col = c("Black","Blue","Red", "Purple", alpha = 70))
```

###Graph:
```{r}
Gh18_MDS_I3 = as.data.frame(ghsoilmds[["points"]])
Gh18_Ord_I3 = bind_cols(Gh18_data_I3[1:31],Gh18_MDS_I3)

Gh18_MDS_I3 = ggplot(data = Gh18_Ord_I3, aes(MDS1,MDS2))+
    geom_point(aes(color=Clade,shape=Soil_type.x))+
    #stat_ellipse(aes(fill = Clade_Soil_Time), geom="polygon",level=0.95,alpha=0.2) +
    theme_bw()
Gh18_MDS_I3  

#PCoA Plot:
Gh18_PCoA_I3 = subset(PCoA_GH, Iteration == 3 & Selection.x != "S")
Gh18_Plot_I3 = ggplot(data = Gh18_PCoA_I3 , aes(Axis.1,Axis.2))+
    geom_point(aes(color=Clade,shape=Soil_type.x))+
    #stat_ellipse(aes(fill = Clade_Soil_Time), geom="polygon",level=0.95,alpha=0.2) +
    labs(x = "PCoA 1 (13.4%)", y = "PCoA 2 (8.4%)")+
    theme_bw()
Gh18_Plot_I3

```

###model for iteration 3
```{r}
fit_I3 = adonis(Gh18_data_I3[32:1694] ~ Soil_type.x * Clade * Selection.x, data = Gh18_data_I3, permutation = 999)
fit_I3

#Pairwise Permanova
library(pairwiseAdonis)
pairwise.adonis2(Gh18_data_I3[32:1694] ~ Clade, data = Gh18_data_I3)

```

```{r}
pairwise.adonis2(Gh18_data_I1[31:1694] ~ Clade, data = Gh18_data_I1)

```

```{r}
pairwise.adonis2(Gh18_data_I5[29:1695] ~ Clade, data = Gh18_data_I5)
```


So, we do see microbial communities begin to diverge relative to each other across iterations 1->3->5 

Specifically, we see that Soil_type.x and Clade are starting to make a difference in terms of fixed effect but there's not interaction effect between soil and clade...

