---
title: "Beta_diversity and Yield"
output: html_notebook
---


```{r}
library(dplyr)
library(vegan)
library(tidyr)
library(ggplot2)
library(readr)
library(reshape)
library(ape)
library(RColorBrewer)
```

*Data quality control*
```{r}

Gh18_data = read.csv("Gh18_ASV_data.csv", header = T)
Gh18_data1 = cbind(Gh18_data,total=rowSums(Gh18_data[31:1695]))
#Quality control step:
Gh18_data2 = subset(Gh18_data1,total>2000)

```


*Making new data frame*
```{r}
Gh18_data3 = decostand(Gh18_data2[32:1695], method = "total")
Gh18_data3= bind_cols(Gh18_data2[1:31],Gh18_data3)
```


## Hypothesis testing:

We want to ask the question is there connection between Yield and Microbiome composition? And also make a Ordination plot for this...

## **Andigenum** 
```{r}
Gh18_data3_A = subset(Gh18_data3,Selection.x != "S" & Clade == "A")

set.seed(2002)
dist <- vegdist(Gh18_data3_A[32:1694],method = "bray")
library(ape)
PCOA <- pcoa(dist)
eigen_plot <- barplot(PCOA$values$Relative_eig[1:10])

```


#### **Ordination Plot Andigenum**
```{r}
Gh18_data3_A = subset(Gh18_data3,Selection.x != "S" & Clade == "A")

PCOA <- pcoa(dist, correction = "cailliez")
biplot.pcoa(PCOA)
PCoA_GH <- cbind(Gh18_data3_A[1:31],PCOA[["vectors"]])
PCoA_GH$Iteration = as.factor(PCoA_GH$Iteration)

Gh18_PCoA = ggplot(data = PCoA_GH, aes(Axis.1,Axis.2))+
    geom_point(aes(color=Response,shape=Selection.x))+
    stat_ellipse(aes(fill = Soil_type.x), geom="polygon",level=0.95,alpha=0.2)+
    scale_colour_gradient(low = "blue", high="red")+
    theme_bw()
Gh18_PCoA

```
Color is hard to see in terms of response...maybe break it down by iteration:


```{r}
Gh18_data3_PN5 = subset(Gh18_data3,Selection.x != "S" & Clade == "A" & Iteration == 5 & X != 666)

#Removed sample 666 because it is an non-influential outlier

set.seed(2003)
dist <- vegdist(Gh18_data3_PN5[32:1695],method = "bray")
library(ape)
PCOA <- pcoa(dist)
eigen_plot <- barplot(PCOA$values$Relative_eig[1:10])

PCOA <- pcoa(dist, correction = "cailliez")
biplot.pcoa(PCOA)
PCoA_GH <- cbind(Gh18_data3_PN5[1:31],PCOA[["vectors"]])
PCoA_GH$Iteration = as.factor(PCoA_GH$Iteration)

Gh18_PCoA5 = ggplot(data = PCoA_GH, aes(Axis.1,Axis.2))+
  geom_point(aes(color=Response,shape=Selection.x, size=Soil_type.x))+
  stat_ellipse(aes(fill = Soil_type.x), geom="polygon",level=0.95,alpha=0.2)+
  scale_colour_gradient(low = "yellow", high="dark green")+
  theme_bw()
Gh18_PCoA5

```


Definitely see a gradient of response and microbial community at iteration 5 for Andigenum. See trend where Winterfell soils generally had a more positive response with Andigenum potatoes although this is weak statistical evidence. 

**Andigenum and Response (Table 4A for Ch3)**
```{r}
set.seed(2010)
fit_I5 = adonis(Gh18_data3_PN5[32:1695] ~ Soil_type.x * Response * Selection.x, data = Gh18_data3_PN5, permutations = 999)
fit_I5
```
Soil Type and Response effect was present @ iteration 5. 

Response is a "correlation" 


## **Chiletanum Figure 3B for Ch 3**
```{r}
Gh18_data3_C = subset(Gh18_data3, Soil_type.x != "None" & Clade == "C")

set.seed(2003)
dist <- vegdist(Gh18_data3_C[32:1695],method = "bray")
library(ape)
PCOA <- pcoa(dist)
eigen_plot <- barplot(PCOA$values$Relative_eig[1:10])

PCOA <- pcoa(dist, correction = "cailliez")
biplot.pcoa(PCOA)
PCoA_GH <- cbind(Gh18_data3_C[1:31],PCOA[["vectors"]])
PCoA_GH$Iteration = as.factor(PCoA_GH$Iteration)

Gh18_PCoAC = ggplot(data = PCoA_GH, aes(Axis.1,Axis.2))+
    geom_point(aes(color=Response,shape=Soil_type.x))+
    stat_ellipse(aes(fill = Iteration), geom="polygon",level=0.95,alpha=0.2)+
    scale_colour_gradient(low = "yellow", high="dark green")+
    theme_bw()
Gh18_PCoAC

```


```{r}
Gh18_data3_C5 = subset(Gh18_data3,Selection.x != "S" & Clade == "C" & Iteration == 5 & X != 685 & X != 677)

set.seed(2006)
dist <- vegdist(Gh18_data3_C5[32:1695],method = "bray")
library(ape)
PCOA <- pcoa(dist)

PCOA <- pcoa(dist, correction = "cailliez")
biplot.pcoa(PCOA)
PCoA_GH <- cbind(Gh18_data3_C5[1:31],PCOA[["vectors"]])
PCoA_GH$Iteration = as.factor(PCoA_GH$Iteration)

Gh18_PCoA_C5 = ggplot(data = PCoA_GH, aes(Axis.1,Axis.2))+
    geom_point(aes(color=Response,shape=Selection.x,size=Soil_type.x))+
    #geom_text(aes(label=X), size = 4)+
    stat_ellipse(aes(fill = Soil_type.x), geom="polygon",level=0.95,alpha=0.2)+
    scale_colour_gradient(low = "yellow", high="dark green")+
    theme_bw()
Gh18_PCoA_C5

```



```{r}
set.seed(2011)
fit_I5 = adonis(Gh18_data3_C5[32:1695] ~ Soil_type.x * Response * Selection.x, data = Gh18_data3_C5, permutations = 999)
fit_I5
```
Rick: Communities correlate with response maybe different based on soil type. 

## *Modern*
```{r}
Gh18_data3_M = subset(Gh18_data3,Selection.x != "S" & Clade == "M")

set.seed(2010)
dist <- vegdist(Gh18_data3_M[32:1695],method = "bray")
library(ape)
PCOA <- pcoa(dist)
eigen_plot <- barplot(PCOA$values$Relative_eig[1:10])

PCOA <- pcoa(dist, correction = "cailliez")
biplot.pcoa(PCOA)
PCoA_GH <- cbind(Gh18_data3_M[1:31],PCOA[["vectors"]])
PCoA_GH$Iteration = as.factor(PCoA_GH$Iteration)

Gh18_PCoAM = ggplot(data = PCoA_GH, aes(Axis.1,Axis.2))+
    geom_point(aes(color=Response,shape=Soil_type.x,size=Soil_type.x))+
    stat_ellipse(aes(fill = Soil_type.x), geom="polygon",level=0.95,alpha=0.2)+
    scale_colour_gradient(low = "yellow", high="dark green")+
    theme_bw()
Gh18_PCoAM
```



```{r}
Gh18_data3_M5 = subset(Gh18_data3,Selection.x != "S" & Clade == "M" & Iteration == 5 & X != 726 & X != 738)

set.seed(2006)
dist <- vegdist(Gh18_data3_M5[32:1695],method = "bray")
library(ape)
PCOA <- pcoa(dist)

PCOA <- pcoa(dist, correction = "cailliez")
biplot.pcoa(PCOA)
PCoA_GH <- cbind(Gh18_data3_M5[1:31],PCOA[["vectors"]])
PCoA_GH$Iteration = as.factor(PCoA_GH$Iteration)

Gh18_PCoA_M5 = ggplot(data = PCoA_GH, aes(Axis.1,Axis.2))+
    geom_point(aes(color=Response,shape=Selection.x), size = 7)+
    #geom_text(aes(label = X), size = 4)+
    stat_ellipse(aes(fill = Soil_type.x), geom="polygon",level=0.95,alpha=0.2)+
    scale_colour_gradient(low = "yellow", high="dark green")+
    theme_bw()
Gh18_PCoA_M5
```

**Rick's Comments:**

> Rick: Generally a single bad state and mutiple good state. Surface conture plot in vegan ordisurf. Takes two dimensional ord and then makes topo map of the ordination space. Look at hills and toughs. THings may not be linear for ordination. When they are bad similar...but when they are good...

>Note on outliers, maybe think about removing those points...



```{r}
set.seed(2012)
Gh18_data3_M5P = subset(Gh18_data3_M5, Selection.x == "P")
fit_I5_P = adonis(Gh18_data3_M5P[32:1695] ~ Soil_type.x * Response, data = Gh18_data3_M5P, permutations = 999)
fit_I5_P

set.seed(2013)
Gh18_data3_M5N = subset(Gh18_data3_M5, Selection.x == "N")
fit_I5_N = adonis(Gh18_data3_M5N[32:1695] ~ Soil_type.x * Response, data = Gh18_data3_M5N, permutations = 999)
fit_I5_N
```

Only see Soil_type effect after removing outliers. 

This makes sense in that in Moderns its symbiotic relationship with a few key rare microbial taxa that make yield boost observed. DESeq2 data also conveys this evidence...

*Rick's Comments:*

>Different between soil types leading to different soil communities that are responsive to potato. Multiple ways to skin a cat. There was not visible disease but may some parasites...

>Sidenote: predict yield --> microbiome data. Lasso regression. Last step for this analysis. What taxa are underlying positive response. Better than DESeq because it's linear and you can put in your predictors. Field yield data, for common scab more reproducible at phyla and order taxa but maybe different b/c greenhouse data more consistent. Do for iteration 3 and 5. you can build a complex linear model and shrinkage of variables...


####Deeper dive into moderns: 

Based on what we saw for the moderns. We are interested in examining MC in more detail for selection and soil type. 

```{r}
Gh18_data3_MH5 = subset(Gh18_data3,Selection.x != "S" & Clade == "M" & Iteration == 5 & Soil_type.x == "Hancock")

set.seed(2013)
fit_IH5 = adonis(Gh18_data3_MH5[32:1695] ~ Response * Selection.x, data = Gh18_data3_MH5, permutations = 999)
fit_IH5

Gh18_data3_MP5 = subset(Gh18_data3,Selection.x != "S" & Clade == "M" & Iteration == 5 & Selection.x == "P")

set.seed(2014)
fit_IP5 = adonis(Gh18_data3_MP5[32:1695] ~ Response * Soil_type.x, data = Gh18_data3_MP5, permutations = 999)
fit_IP5

```
Response:Soil_type.x interaction effect tied to the microbial community $p-value$ = 0.001

Positively selected soils tied to shaping the microbial community depending on soil type. 

It looks like two different microbial states existing here states for positively selected microbial communities...Its not as nice and neat as what I thought ofc....


##Winterfell soils

```{r}
Gh18_data3_MH5 = subset(Gh18_data3,Selection.x != "S" & Clade == "M" & Iteration == 5 & Soil_type.x == "Winterfell")

set.seed(2030)
fit_IH5 = adonis(Gh18_data3_MH5[32:1695] ~ Response * Selection.x, data = Gh18_data3_MH5, permutations = 999)
fit_IH5

Gh18_data3_MP5 = subset(Gh18_data3,Selection.x != "S" & Clade == "M" & Iteration == 5 & Selection.x == "P")

Gh18_data3_M5 = subset(Gh18_data3,Selection.x != "S" & Clade == "M" & Iteration == 5 & X != 726 & X != 738)


set.seed(2031)
fit_IP5 = adonis(Gh18_data3_M5[32:1695] ~ Response * Soil_type.x * Selection.x, data = Gh18_data3_M5, permutations = 999)
fit_IP5
```

