---
title: "Phyloseq_analysis"
output: html_notebook
---


```{r setup, include=FALSE}
#library(seqinr)
library(reshape)
library(plyr)
library(dplyr)
library(plotly)
library(dplyr)
library(tidyr)
library(ggplot2)
library(vegan)
library(RColorBrewer)
library(tidyverse)
library(gridExtra)
#library(limma)
library(maps)
library(reshape2)
library(broom)
library(ggpubr)
library(phyloseq)
library(DESeq2)
library(stats4)
library(minpack.lm)
library(Hmisc)
library(tidytext)
library(stringr)
library(ggrepel)
library(ggExtra)
```

```{r}
taxa_table = read_tsv('merged-taxonomy.tsv')
ASV_table = read_tsv('16S-merged-table.txt')

#somehow need to make #OTU_ID column into rownames:
ASV_table1 = ASV_table[,-1] #make a new dataframe
rownames(ASV_table1) = ASV_table$`#OTU_ID` #assign rownames to OTU_ID from original dataframe
#replace NAs in ASV table to zero:
ASV_table1[is.na(ASV_table1)] = 0

#problem ASV_table1 dataframe is a tibble, therefore need to reset it to a dataframe:
ASV_table1 = as.data.frame(ASV_table1)

#Preparing ASV table
otu_us <- ASV_table1[!grepl("Mitochondria", ASV_table$taxonomy),]
otu_us <- otu_us[!grepl("Chloroplast", otu_us$taxonomy),]

#Preparing TAXA
tax_otu <- otu_us['taxonomy']
tax_otu <- tax_otu %>%
  separate(taxonomy, into=c("Kingdom", "Phylum", "Class", 
                            "Order", "Family", "Genus", "Species"), sep="; ", remove=F)
tax_otu[2:8] <- lapply(tax_otu[2:8], function(x) gsub(".*__", "", x))

#Preparing OTU table again
otu_us['taxonomy'] <- NULL

#Preparing mapping file: 
map_table = read.csv("metadata3.csv", header = T)
map_table1 = map_table[,-9] #make a new dataframe
rownames(map_table1) = map_table$Submission_ID #assign rownames to OTU_ID from original dataframe
map_table1[1] <- NULL
```



```{r}
#Taxonomic distribution 
tax_v1 <- tax_otu
# creating column where last taxonomic identifier is attached to the OTU id 
features <- c(sprintf("OTU%05d", seq(1,10614)),"label")
tax_v1$otu_id <- features
tax_v1 <- tax_v1 %>% select(otu_id, everything())

lastValue <- function(x) tail(x[!is.na(x)], 1)
last_taxons<- apply(tax_v1, 1, lastValue)
tax_v1$last_taxon <- last_taxons
tax_v1$final_names <- paste(tax_v1$otu_id, tax_v1$last_taxon, sep='-')

tax_v1$otu <- rownames(tax_v1)
```

#Generate Phyloseq Object:
```{r}
taxonomy_phyloseq<- tax_v1[rownames(tax_v1) %in% rownames(otu_us),]

#Make phyloseq object:
OTU <- otu_table(otu_us, taxa_are_rows=T)
TAX <- tax_table(as.matrix(taxonomy_phyloseq))
metadat <- sample_data(map_table1,errorIfNULL=TRUE)
taxa_names(OTU)
physeq <- merge_phyloseq(phyloseq(OTU, TAX), metadat)

#physeq = subset_samples(physeq, Soil_type.x != "Sterile")

```

#Quality Control:
```{r}
#Filter samples that aren't relevant for analysis:
physeq_filtered = subset_samples(physeq, Soil_inoculation != "Sterile" & Response != "NA")
physeq_filtered <- subset_taxa(physeq_filtered, !is.na(Phylum) & !Phylum %in% c( "uncharacterized"))


#Quality control steps of OTU table:
physeq_trimmed <- prune_taxa(taxa_sums(physeq_filtered) > 0, physeq_filtered) #Remove taxa where total OTU count aka row count is greater than zero.
physeq_trimmed <- prune_samples(sample_sums(physeq_trimmed)>=2000,physeq_trimmed) #Remove samples that contain less that 2000 reads, now we only have 271 samples:


# Re-order the columns
tax_table(physeq_trimmed) <- tax_table(physeq_trimmed)[,c(3:12,1,2)]
rank.names(physeq_trimmed)


#physeq_phyla = tax_glom(physeq_trimmed,taxrank=rank_names(physeq_trimmed)[2])
#ps.relabund = transform_sample_counts(physeq_trimmed, function(x) x/sum(x))

```

#Export phyloseq object:
```{r}

#For differential analysis: 
#saveRDS(physeq_trimmed, file="~/Library/CloudStorage/Box-Box/Lankau_Lab/Greenhouse_Exmpt_2018/16S/Differential_analysis/ps_merged.RDS")

#For alpha diversity analysis: 
saveRDS(physeq_trimmed, file="~/Library/CloudStorage/Box-Box/Lankau_Lab/Greenhouse_Exmpt_2018/16S/ps_merged.RDS")
```


Prevalance filtering: 
```{r}
prevdf = apply(X = otu_table(physeq_trimmed),
               MARGIN = ifelse(taxa_are_rows(physeq_trimmed), yes = 1, no = 2),
               FUN = function(x){sum(x > 0)})

# Add taxonomy and total read counts to this data.frame
prevdf = data.frame(Prevalence = prevdf,
                    TotalAbundance = taxa_sums(physeq_trimmed),
                    tax_table(physeq_trimmed))

prevdf1 = subset(prevdf, otu %in% get_taxa_unique(physeq_trimmed, "otu"))

ggplot(prevdf1, aes(TotalAbundance, Prevalence / nsamples(physeq_trimmed),color=Phylum)) +
  # Include a guess for parameter hline (5%)...
  geom_hline(yintercept = 0.05, alpha = 0.5, linetype = 2) +  
  geom_point(size = 2, alpha = 0.7) +
  scale_x_log10() +  
  xlab("Total Abundance") + 
  ylab("Prevalence [Frac. Samples]") +
  facet_wrap(~Phylum) + 
  theme(legend.position="none")


prevalenceThreshold = 2
prevalenceThreshold

# Execute prevalence filter, using `prune_taxa()` function
keepTaxa = rownames(prevdf1)[(prevdf1$Prevalence >= 2)]
ps1 = prune_taxa(keepTaxa, physeq_trimmed)
#saveRDS(ps1,"ps1.RDS")

#saveRDS(ps1, file="~/Library/CloudStorage/Box-Box/Lankau_Lab/Greenhouse_Exmpt_2018/16S/Differential_analysis/ps1.RDS")

#Keep taxa where prevalance is greater than 2

```



#Stacked Barchart of Phyloseq object:

```{r}
#Aggregate at the Phyla Level: 
physeq_trimmed_i1 = subset_samples(ps1, Iteration == 1)
physeq_trimmed_i5 = subset_samples(ps1, Iteration == 5)


physeq_phyla_i1 = tax_glom(physeq_trimmed_i1, taxrank = rank_names(physeq_trimmed_i1)[2])
physeq_phyla_i5 = tax_glom(physeq_trimmed_i5, taxrank = rank_names(physeq_trimmed_i5)[2])

merged_clade_i1 = merge_samples(physeq_phyla_i1,"Clade")
merged_clade_i5 = merge_samples(physeq_phyla_i5,"Clade")

```

Warning "NA introduced by coercion" we can ignore since it seems to be only the metadata that is effect by this....


relativize sample counts: 
```{r}
ps.relabundPhyla_i1 = transform_sample_counts(merged_clade_i1, function(x) x/sum(x))
ps.relabundPhyla_i5 = transform_sample_counts(merged_clade_i5, function(x) x/sum(x))
```


Plot stacked bar chart:
```{r}
plot_bar(ps.relabundPhyla_i1, fill = "Phylum")
plot_bar(ps.relabundPhyla_i5, fill = "Phylum")

```

Color scheme isn't super great...let's change this to something a bit more discrete:


```{r}
library(RColorBrewer)

nb.cols <- 36
palette_colors = colorRampPalette(brewer.pal(8,"Set1"))(nb.cols)

Clade_StackPlot_i1 = plot_bar(ps.relabundPhyla_i1, fill = "Phylum") + 
  scale_fill_manual(values = palette_colors)+
  xlab(expression(Potato~Clade))+
  scale_x_discrete(labels=c("Andigenum","Chiletanum","Modern"))+
  ylab(expression(Relative~Abundance))+
  theme_bw()
Clade_StackPlot_i1
```


```{r}
nb.cols <- 36
palette_colors = colorRampPalette(brewer.pal(8,"Set1"))(nb.cols)

Clade_StackPlot_i5 = plot_bar(ps.relabundPhyla_i5, fill = "Phylum") + 
  scale_fill_manual(values = palette_colors)+
  xlab(expression(Potato~Clade))+
  scale_x_discrete(labels=c("Andigenum","Chiletanum","Modern"))+
  ylab(expression(Relative~Abundance))+
  theme_bw()
Clade_StackPlot_i5
```


**Relative abundance for + vs - responsive Plants for each Clade (Ch 3):**

```{r}
#Aggregate at the Phyla Level: 
physeq_trimmed_P = subset_samples(ps1, Type_Response == "Positive")
physeq_trimmed_N = subset_samples(ps1, Type_Response == "Negative")

physeq_phyla_P = tax_glom(physeq_trimmed_P, taxrank = rank_names(physeq_trimmed_P)[2])
physeq_phyla_N = tax_glom(physeq_trimmed_N, taxrank = rank_names(physeq_trimmed_N)[2])

merged_clade_P = merge_samples(physeq_phyla_P,"Clade")
merged_clade_N = merge_samples(physeq_phyla_N,"Clade")
```

```{r}
ps.relabundPhyla_P = transform_sample_counts(merged_clade_P, function(x) x/sum(x))
ps.relabundPhyla_N = transform_sample_counts(merged_clade_N, function(x) x/sum(x))
```


```{r}
nb.cols <- 36
palette_colors = colorRampPalette(brewer.pal(8,"Set1"))(nb.cols)

Clade_StackPlot_P = plot_bar(ps.relabundPhyla_P, fill = "Phylum") + 
  scale_fill_manual(values = palette_colors)+
  xlab(expression(Potato~Clade))+
  scale_x_discrete(labels=c("Andigenum","Chiletanum","Modern"))+
  ylab(expression(Relative~Abundance))+
  theme_bw()
Clade_StackPlot_P
```


```{r}
nb.cols <- 36
palette_colors = colorRampPalette(brewer.pal(8,"Set1"))(nb.cols)

Clade_StackPlot_N = plot_bar(ps.relabundPhyla_N, fill = "Phylum") + 
  scale_fill_manual(values = palette_colors)+
  xlab(expression(Potato~Clade))+
  scale_x_discrete(labels=c("Andigenum","Chiletanum","Modern"))+
  ylab(expression(Relative~Abundance))+
  theme_bw()
Clade_StackPlot_N
```





