---
title: "Generating Stacked BarPlot for Contrasts DESeq2: Figure 4"
output: html_notebook
---

We were able to come up with the correct code to generate stacked barplots (See 'DESeq2_clade_1202026.Rmd'). 

We will now Generate figures for phyla level differences and try to make publication ready graphs. Currently the manuscript for figures 3,4,5,6, and 7 aren't really ideal for publication. 

We want to generate a figure that is similar to this: https://doi.org/10.1038/s41598-020-72197-y (See Figure 2A)

We want to generate similar plot to Figure 2A. 

```{r}
library(DESeq2)
library(ggplot2)
library(dplyr)
library(tidyr)
library(ggpubr)
library(RColorBrewer)
library(phyloseq)
```

Quality control of Phyloseq object: 
```{r}
#Load R environment with our PhyloSeq object:
ps1 = readRDS("/Users/maxmiao/Library/CloudStorage/Box-Box/Lankau_Lab/Ch3_Greenhouse_Exmpt_2018/Phyloseq_analysis/ps1_DESeq2_02232026.RDS")
ps1_filt <- prune_samples(sample_sums(ps1) >= 8000, ps1)

#Aggregate counts based on taxa rank (in this case Phyla)
ps1_df = tax_glom(ps1_filt,taxrank=rank_names(ps1_filt)[5])
ps1_df = subset_taxa(ps1_df, Phylum != "NA")
ps1_df = subset_samples(ps1_df, sample_sums(ps1_df) != 0)

#Now we want to separate based on Time. (Week 3, Week 9, Week 15)
ps1_df_W3 = subset_samples(ps1_df, Iteration == '1')
ps1_df_W9 = subset_samples(ps1_df, Iteration == '3')
ps1_df_W15 = subset_samples(ps1_df, Iteration == '5')

#Convert Phyoseq objects into DESeq2 objects: 
dds1_W3= phyloseq_to_deseq2(ps1_df_W3, ~ Clade)
dds1_W9= phyloseq_to_deseq2(ps1_df_W9, ~ Clade)
dds1_W15= phyloseq_to_deseq2(ps1_df_W15, ~ Clade)

```
#Normalization step for our samples: 


Normalization using 'esimateSizeFactor' function: Accounts for variation in total read counts between samples so that counts can be comparable across samples without being biased by sequencing depth. We specify type = "poscounts" due to inflated zeros for 16S data. Yields relative scaling and is ideal of compositional data like 16S amplicon sequencing. 


```{r}


design(dds1_W3) <- ~ 1 + Clade
design(dds1_W9) <- ~ 1 + Clade
design(dds1_W15) <- ~ 1 + Clade

dds2_W3 = estimateSizeFactors(dds1_W3, type = "poscounts") #poscounts is better for 16S data
dds2_W9 = estimateSizeFactors(dds1_W9, type = "poscounts") 
dds2_W15 = estimateSizeFactors(dds1_W15, type = "poscounts") 


#Re-rerun DESeq2 on subset data after data normalization: 
dds3_W3 = DESeq(dds2_W3, fitType="parametric")
dds3_W9 = DESeq(dds2_W9, fitType="parametric")
dds3_W15 = DESeq(dds2_W15, fitType="parametric")

```

Now we will generate contrasts: 
```{r}
#Calculating coefficient vectors for DESeq2 output for each potato Clade
mod_mat = model.matrix(design(dds3_W3), colData(dds3_W3))

North_Am <- colMeans(mod_mat[dds3_W3$Clade == "M", ])
Chiletanum <- colMeans(mod_mat[dds3_W3$Clade == "C", ])
Andigenum <- colMeans(mod_mat[dds3_W3$Clade == "A", ])

#And we can now define any contrasts we want:

##For week 3 contrasts
res1_Mod_Chil <- results(dds3_W3, contrast = North_Am - Chiletanum)
res1_Mod_And <- results(dds3_W3, contrast = North_Am - Andigenum)
res1_Chil_And <- results(dds3_W3, contrast = Chiletanum - Andigenum)

#Glue phyla column to DESeq2 object in our results dataframe based on rowname in both objects!


```



Prepare Data for Plotting using ggplot:
```{r}

res_Test_Phyla = cbind(as(res1_Mod_And, "data.frame"), as(tax_table(ps1_df_W3)[rownames(res1_Mod_And), ], "matrix"))

# Add significance indicator
res_df <- res_Test_Phyla %>%
  mutate(Significant = ifelse(padj < 0.05, "Yes", "No"), 
         Direction = ifelse(log2FoldChange > 0, "Increased", "Decreased"))

res_df = filter(res_df, res_df$Significant == "Yes")

  
# Create color palette
  color_palette <- c("Increased" = "#e31a1c", "Decreased" = "#1f78b4")
  
# Create the horizontal barplot
p1_1 <- ggplot(res_df, aes(x = log2FoldChange, y = Family, fill = Direction)) +
  geom_bar(stat = "identity", width = 0.7) +
  scale_fill_manual(values = color_palette, 
                      name = "Direction of Change",
                      labels = c("Decreased", "Increased")) +
    # Add vertical line at zero
    geom_vline(xintercept = 0, linetype = "dashed", color = "gray40", linewidth = 0.5) +
  #Add significance indicators
   geom_point(data = subset(res_df, Significant == "Yes"),
              aes(x = ifelse(log2FoldChange > 0, 
                              log2FoldChange + max(abs(log2FoldChange))*0.05,
                              log2FoldChange - max(abs(log2FoldChange))*0.05)),
             shape = 8, size = 2, color = "black", show.legend = FALSE) +
    labs(
    title = "N. America - Andigenum",
    x = expression(bold("Log"[2]*" Fold Change")),
    y = "Family") +
  
  theme_minimal(base_size = 12) +
    theme(
      plot.title = element_text(face = "bold", hjust = 0.5, size = 12),
      plot.subtitle = element_text(hjust = 0.5, color = "gray40", size = 14),
      axis.title = element_text(face = "bold"),
      axis.text.y = element_text(size = 11),
      axis.text.x = element_text(size = 11),
      legend.position = "bottom",
      legend.title = element_text(face = "bold"),
      panel.grid.major.y = element_blank(),
      panel.grid.minor.y = element_blank(),
      panel.grid.major.x = element_line(color = "gray90"),
      panel.grid.minor.x = element_blank(),
      plot.caption = element_text(color = "gray50", size = 9)
    )

p1_1
```
Nice! We were able to get a nice graph!! for Iteration 1 (After 3 months) NA - Chiletanum variety.
As we can see despite the changes in abundance across multiple phyla none were significantly significant at Iteration 1. 


Let's continue for our other contrasts for iteration 1: NA-Andigenum and Chiletanum - Andigenum. 

```{r}
#Note we saved our results in lines 94 as: 
#res1_Mod_Chil <- results(dds3_W3, contrast = North_Am - Chiletanum)
#res1_Chil_And <- results(dds3_W3, contrast = Chiletanum - Andigenum)

#Glue phyla column to DESeq2 object in our results dataframe based on rowname in both objects!
res2_NA_Chil_Phyla = cbind(as(res1_Mod_Chil, "data.frame"), as(tax_table(ps1_df_W3)[rownames(res1_Mod_Chil), ], "matrix"))

res2_Chil_And_Phyla = cbind(as(res1_Chil_And, "data.frame"), as(tax_table(ps1_df_W3)[rownames(res1_Chil_And), ], "matrix"))

```



Prepare Data for Plotting using ggplot (Iteration 1 Northern American vs Chiletanum):
```{r}

# Add significance indicator
res_df <- res2_NA_Chil_Phyla %>%
  mutate(Significant = ifelse(padj < 0.05, "Yes", "No"), 
         Direction = ifelse(log2FoldChange > 0, "Increased", "Decreased"))
res_df = filter(res_df, res_df$Significant == "Yes")

# Create color palette
  color_palette <- c("Increased" = "#e31a1c", "Decreased" = "#1f78b4")
  
# Create the horizontal barplot
p1_2 <- ggplot(res_df, aes(x = log2FoldChange, y = Family, fill = Direction)) +
  geom_bar(stat = "identity", width = 0.7) +
  scale_fill_manual(values = color_palette, 
                      name = "Direction of Change",
                      labels = c("Decreased", "Increased")) +
    # Add vertical line at zero
    geom_vline(xintercept = 0, linetype = "dashed", color = "gray40", linewidth = 0.5) +
  #Add significance indicators
   geom_point(data = subset(res_df, Significant == "Yes"),
              aes(x = ifelse(log2FoldChange > 0, 
                              log2FoldChange + max(abs(log2FoldChange))*0.05,
                              log2FoldChange - max(abs(log2FoldChange))*0.05)),
             shape = 8, size = 2, color = "black", show.legend = FALSE) +
  
  labs(
    title = "N. America - Chiletanum",
    x = expression(bold("Log"[2]*" Fold Change")),
    y = "Family"
  ) +
  
  theme_minimal(base_size = 12) +
    theme(
      plot.title = element_text(face = "bold", hjust = 0.5, size = 12),
      plot.subtitle = element_text(hjust = 0.5, color = "gray40", size = 14),
      axis.title = element_text(face = "bold"),
      axis.text.y = element_text(size = 11),
      axis.text.x = element_text(size = 11),
      legend.position = "bottom",
      legend.title = element_text(face = "bold"),
      panel.grid.major.y = element_blank(),
      panel.grid.minor.y = element_blank(),
      panel.grid.major.x = element_line(color = "gray90"),
      panel.grid.minor.x = element_blank(),
      plot.caption = element_text(color = "gray50", size = 9)
    )

p1_2 
```

Contrasts between Chiletanum vs Andigenum at iteration 1 (3 months)
```{r}
# Add significance indicator
  res_df <- res2_Chil_And_Phyla %>%
    mutate(Significant = ifelse(padj < 0.05, "Yes", "No"), Direction = ifelse(log2FoldChange > 0, "Increased", "Decreased")
    )
res_df = filter(res_df, res_df$Significant == "Yes")

  
# Create color palette
  color_palette <- c("Increased" = "#e31a1c", "Decreased" = "#1f78b4")
  
# Create the horizontal barplot
p1_3 <- ggplot(res_df, aes(x = log2FoldChange, y = Family, fill = Direction)) +
  geom_bar(stat = "identity", width = 0.7) +
  scale_fill_manual(values = color_palette, 
                      name = "Direction of Change",
                      labels = c("Decreased", "Increased")) +
    # Add vertical line at zero
    geom_vline(xintercept = 0, linetype = "dashed", color = "gray40", linewidth = 0.5) +
  #Add significance indicators
   geom_point(data = subset(res_df, Significant == "Yes"),
              aes(x = ifelse(log2FoldChange > 0, 
                              log2FoldChange + max(abs(log2FoldChange))*0.05,
                              log2FoldChange - max(abs(log2FoldChange))*0.05)),
             shape = 8, size = 2, color = "black", show.legend = FALSE) +
  
    labs(
    title = "Chiletanum - Andigenum",
    x = expression(bold("Log"[2]*" Fold Change")),
    y = "Family"
  ) +
  
  theme_minimal(base_size = 12) +
    theme(
      plot.title = element_text(face = "bold", hjust = 0.5, size = 12),
      plot.subtitle = element_text(hjust = 0.5, color = "gray40", size = 14),
      axis.title = element_text(face = "bold"),
      axis.text.y = element_text(size = 11),
      axis.text.x = element_text(size = 11),
      legend.position = "bottom",
      legend.title = element_text(face = "bold"),
      panel.grid.major.y = element_blank(),
      panel.grid.minor.y = element_blank(),
      panel.grid.major.x = element_line(color = "gray90"),
      panel.grid.minor.x = element_blank(),
      plot.caption = element_text(color = "gray50", size = 9)
    )

p1_3
```


***

Let's now do the same but for Iteration 3 (after 9 months)

```{r}
#From line 76 we generated the object: 
dds3_W9 = DESeq(dds2_W9, fitType="parametric")

#Calculating coefficient vectors for DESeq2 output for each potato Clade
mod_mat = model.matrix(design(dds3_W9), colData(dds3_W9))

North_Am <- colMeans(mod_mat[dds3_W9$Clade == "M", ])
Chiletanum <- colMeans(mod_mat[dds3_W9$Clade == "C", ])
Andigenum <- colMeans(mod_mat[dds3_W9$Clade == "A", ])

#And we can now define any contrasts we want:

##For Iteration 3 contrasts
res1_Mod_Chil <- results(dds3_W9, contrast = North_Am - Chiletanum)
res1_Mod_And <- results(dds3_W9, contrast = North_Am - Andigenum)
res1_Chil_And <- results(dds3_W9, contrast = Chiletanum - Andigenum)
```

```{r}
res2_Phyla = cbind(as(res1_Mod_Chil, "data.frame"), as(tax_table(ps1_df_W3)[rownames(res1_Mod_Chil), ], "matrix"))

res_df <- res2_Phyla %>%
    mutate(Significant = ifelse(padj < 0.05, "Yes", "No"), Direction = ifelse(log2FoldChange > 0, "Increased", "Decreased")
    )
res_df = filter(res_df, res_df$Significant == "Yes")

```


Plot out contrast between N.American vs Chiletanum (Iteration 3)
```{r}
# Create color palette
  color_palette <- c("Increased" = "#e31a1c", "Decreased" = "#1f78b4")
  
# Create the horizontal barplot
p3_2 <- ggplot(res_df, aes(x = log2FoldChange, y = Family, fill = Direction)) +
  geom_bar(stat = "identity", width = 0.7) +
  scale_fill_manual(values = color_palette, 
                      name = "Direction of Change",
                      labels = c("Decreased", "Increased")) +
    # Add vertical line at zero
    geom_vline(xintercept = 0, linetype = "dashed", color = "gray40", linewidth = 0.5) +
  #Add significance indicators
   geom_point(data = subset(res_df, Significant == "Yes"),
              aes(x = ifelse(log2FoldChange > 0, 
                              log2FoldChange + max(abs(log2FoldChange))*0.05,
                              log2FoldChange - max(abs(log2FoldChange))*0.05)),
             shape = 8, size = 2, color = "black", show.legend = FALSE) +
  
    labs(
    x = expression(bold("Log"[2]*" Fold Change")),
    y = "Family"
  ) +
  
  theme_minimal(base_size = 12) +
    theme(
      plot.title = element_text(face = "bold", hjust = 0.5, size = 16),
      plot.subtitle = element_text(hjust = 0.5, color = "gray40", size = 12),
      axis.title = element_text(face = "bold"),
      axis.text.y = element_text(size = 11),
      axis.text.x = element_text(size = 11),
      legend.position = "bottom",
      legend.title = element_text(face = "bold"),
      panel.grid.major.y = element_blank(),
      panel.grid.minor.y = element_blank(),
      panel.grid.major.x = element_line(color = "gray90"),
      panel.grid.minor.x = element_blank(),
      plot.caption = element_text(color = "gray50", size = 9)
    )

p3_2
```

North America vs Andigenum Contrast (Iteration 3)
```{r}
res2_Phyla = cbind(as(res1_Mod_And, "data.frame"), as(tax_table(ps1_df_W3)[rownames(res1_Mod_Chil), ], "matrix"))

res_df <- res2_Phyla %>%
    mutate(Significant = ifelse(padj < 0.05, "Yes", "No"), Direction = ifelse(log2FoldChange > 0, "Increased", "Decreased")
    )
res_df = filter(res_df, res_df$Significant == "Yes")

```


```{r}
# Create color palette
  color_palette <- c("Increased" = "#e31a1c", "Decreased" = "#1f78b4")
  
# Create the horizontal barplot
p3_1 <- ggplot(res_df, aes(x = log2FoldChange, y = Family, fill = Direction)) +
  geom_bar(stat = "identity", width = 0.7) +
  scale_fill_manual(values = color_palette, 
                      name = "Direction of Change",
                      labels = c("Decreased", "Increased")) +
    # Add vertical line at zero
    geom_vline(xintercept = 0, linetype = "dashed", color = "gray40", linewidth = 0.5) +
  #Add significance indicators
   geom_point(data = subset(res_df, Significant == "Yes"),
              aes(x = ifelse(log2FoldChange > 0, 
                              log2FoldChange + max(abs(log2FoldChange))*0.05,
                              log2FoldChange - max(abs(log2FoldChange))*0.05)),
             shape = 8, size = 2, color = "black", show.legend = FALSE) +
    
  labs(
    x = expression(bold("Log"[2]*" Fold Change")),
    y = "Family"
  ) +
  
  theme_minimal(base_size = 12) +
    theme(
      plot.title = element_text(face = "bold", hjust = 0.5, size = 16),
      plot.subtitle = element_text(hjust = 0.5, color = "gray40", size = 12),
      axis.title = element_text(face = "bold"),
      axis.text.y = element_text(size = 11),
      axis.text.x = element_text(size = 11),
      legend.position = "bottom",
      legend.title = element_text(face = "bold"),
      panel.grid.major.y = element_blank(),
      panel.grid.minor.y = element_blank(),
      panel.grid.major.x = element_line(color = "gray90"),
      panel.grid.minor.x = element_blank(),
      plot.caption = element_text(color = "gray50", size = 9)
    )

p3_1
```


Chiletanum vs Andigenum Contrast Iteration 3 
```{r}
res2_Phyla = cbind(as(res1_Chil_And, "data.frame"), as(tax_table(ps1_df_W3)[rownames(res1_Chil_And), ], "matrix"))

res_df <- res2_Phyla %>%
    mutate(Significant = ifelse(padj < 0.05, "Yes", "No"), Direction = ifelse(log2FoldChange > 0, "Increased", "Decreased")
    )
res_df = filter(res_df, res_df$Significant == "Yes")

```


```{r}
# Create color palette
  color_palette <- c("Increased" = "#e31a1c", "Decreased" = "#1f78b4")
  
# Create the horizontal barplot
p3_3 <- ggplot(res_df, aes(x = log2FoldChange, y = Family, fill = Direction)) +
  geom_bar(stat = "identity", width = 0.7) +
  scale_fill_manual(values = color_palette, 
                      name = "Direction of Change",
                      labels = c("Decreased", "Increased")) +
    # Add vertical line at zero
    geom_vline(xintercept = 0, linetype = "dashed", color = "gray40", linewidth = 0.5) +
   labs(
    x = expression(bold("Log"[2]*" Fold Change")),
    y = "Family"
  ) +
  #Add significance indicators
   geom_point(data = subset(res_df, Significant == "Yes"),
              aes(x = ifelse(log2FoldChange > 0, 
                              log2FoldChange + max(abs(log2FoldChange))*0.05,
                              log2FoldChange - max(abs(log2FoldChange))*0.05)),
             shape = 8, size = 2, color = "black", show.legend = FALSE) +
  theme_minimal(base_size = 12) +
    theme(
      plot.title = element_text(face = "bold", hjust = 0.5, size = 16),
      plot.subtitle = element_text(hjust = 0.5, color = "gray40", size = 12),
      axis.title = element_text(face = "bold"),
      axis.text.y = element_text(size = 11),
      axis.text.x = element_text(size = 11),
      legend.position = "bottom",
      legend.title = element_text(face = "bold"),
      panel.grid.major.y = element_blank(),
      panel.grid.minor.y = element_blank(),
      panel.grid.major.x = element_line(color = "gray90"),
      panel.grid.minor.x = element_blank(),
      plot.caption = element_text(color = "gray50", size = 9)
    )

p3_3
```
Notice that there were differences in Phyla between Andigenum and Chiletanum (but not with Modern)


***

Let's now do the same but for Iteration 5 (after 15 months)

```{r}
#From line 76 we generated the object: 
dds3_W15 = DESeq(dds2_W15, fitType="parametric")

#Calculating coefficient vectors for DESeq2 output for each potato Clade
mod_mat = model.matrix(design(dds3_W15), colData(dds3_W15))

North_Am <- colMeans(mod_mat[dds3_W15$Clade == "M", ])
Chiletanum <- colMeans(mod_mat[dds3_W15$Clade == "C", ])
Andigenum <- colMeans(mod_mat[dds3_W15$Clade == "A", ])

#And we can now define any contrasts we want:

##For Iteration 5 contrasts
res1_Mod_Chil <- results(dds3_W15, contrast = North_Am - Chiletanum)
res1_Mod_And <- results(dds3_W15, contrast = North_Am - Andigenum)
res1_Chil_And <- results(dds3_W15, contrast = Chiletanum - Andigenum)
```

Contrast between N. American Potatoes vs Chiletanum potatoes 
```{r}
res2_Phyla = cbind(as(res1_Mod_Chil, "data.frame"), as(tax_table(ps1_df_W15)[rownames(res1_Mod_Chil), ], "matrix"))

res_df <- res2_Phyla %>%
    mutate(Significant = ifelse(padj < 0.05, "Yes", "No"), Direction = ifelse(log2FoldChange > 0, "Increased", "Decreased"))
res_df = filter(res_df, res_df$Significant == "Yes")

```



```{r}
# Create color palette
  color_palette <- c("Increased" = "#e31a1c", "Decreased" = "#1f78b4")
  
# Create the horizontal barplot
p5_2 <- ggplot(res_df, aes(x = log2FoldChange, y = Family, fill = Direction)) +
  geom_bar(stat = "identity", width = 0.7) +
  scale_fill_manual(values = color_palette, 
                      name = "Direction of Change",
                      labels = c("Decreased", "Increased")) +
    # Add vertical line at zero
    geom_vline(xintercept = 0, linetype = "dashed", color = "gray40", linewidth = 0.5) +
  #Add significance indicators
   geom_point(data = subset(res_df, Significant == "Yes"),
              aes(x = ifelse(log2FoldChange > 0, 
                              log2FoldChange + max(abs(log2FoldChange))*0.05,
                              log2FoldChange - max(abs(log2FoldChange))*0.05)),
             shape = 8, size = 2, color = "black", show.legend = FALSE) +
  theme_minimal(base_size = 12) +
    theme(
      plot.title = element_text(face = "bold", hjust = 0.5, size = 16),
      plot.subtitle = element_text(hjust = 0.5, color = "gray40", size = 12),
      axis.title = element_text(face = "bold"),
      axis.text.y = element_text(size = 11),
      axis.text.x = element_text(size = 11),
      legend.position = "bottom",
      legend.title = element_text(face = "bold"),
      panel.grid.major.y = element_blank(),
      panel.grid.minor.y = element_blank(),
      panel.grid.major.x = element_line(color = "gray90"),
      panel.grid.minor.x = element_blank(),
      plot.caption = element_text(color = "gray50", size = 9)
    )

p5_2
```

Contrast between N. America vs Andigenum (Iteration 5) across Phyla 
```{r}
res2_Phyla = cbind(as(res1_Mod_And, "data.frame"), as(tax_table(ps1_df_W15)[rownames(res1_Mod_And), ], "matrix"))

res_df <- res2_Phyla %>%
    mutate(Significant = ifelse(padj < 0.05, "Yes", "No"), Direction = ifelse(log2FoldChange > 0, "Increased", "Decreased")
    )

res_df = filter(res_df, res_df$Significant == "Yes")

```



```{r}
# Create color palette
  color_palette <- c("Increased" = "#e31a1c", "Decreased" = "#1f78b4")
  
# Create the horizontal barplot
p5_1 <- ggplot(res_df, aes(x = log2FoldChange, y = Family, fill = Direction)) +
  geom_bar(stat = "identity", width = 0.7) +
  scale_fill_manual(values = color_palette, 
                      name = "Direction of Change",
                      labels = c("Decreased", "Increased")) +
    # Add vertical line at zero
    geom_vline(xintercept = 0, linetype = "dashed", color = "gray40", linewidth = 0.5) +
  #Add significance indicators
   geom_point(data = subset(res_df, Significant == "Yes"),
              aes(x = ifelse(log2FoldChange > 0, 
                              log2FoldChange + max(abs(log2FoldChange))*0.05,
                              log2FoldChange - max(abs(log2FoldChange))*0.05)),
             shape = 8, size = 2, color = "black", show.legend = FALSE) +
  theme_minimal(base_size = 12) +
    theme(
      plot.title = element_text(face = "bold", hjust = 0.5, size = 16),
      plot.subtitle = element_text(hjust = 0.5, color = "gray40", size = 12),
      axis.title = element_text(face = "bold"),
      axis.text.y = element_text(size = 11),
      axis.text.x = element_text(size = 11),
      legend.position = "bottom",
      legend.title = element_text(face = "bold"),
      panel.grid.major.y = element_blank(),
      panel.grid.minor.y = element_blank(),
      panel.grid.major.x = element_line(color = "gray90"),
      panel.grid.minor.x = element_blank(),
      plot.caption = element_text(color = "gray50", size = 9)
    )

p5_1
```


Contrast between Chiletanum and Andigenum (Across Phyla) at Iteration 5
```{r}
res2_Phyla = cbind(as(res1_Chil_And, "data.frame"), as(tax_table(ps1_df_W3)[rownames(res1_Chil_And), ], "matrix"))

res_df <- res2_Phyla %>%
    mutate(Significant = ifelse(padj < 0.05, "Yes", "No"), Direction = ifelse(log2FoldChange > 0, "Increased", "Decreased")
    )

res_df = filter(res_df, res_df$Significant == "Yes")

```



```{r}
# Create color palette
  color_palette <- c("Increased" = "#e31a1c", "Decreased" = "#1f78b4")
  
# Create the horizontal barplot
p5_3 <- ggplot(res_df, aes(x = log2FoldChange, y = Family, fill = Direction)) +
  geom_bar(stat = "identity", width = 0.7) +
  scale_fill_manual(values = color_palette, 
                      name = "Direction of Change",
                      labels = c("Decreased", "Increased")) +
    # Add vertical line at zero
    geom_vline(xintercept = 0, linetype = "dashed", color = "gray40", linewidth = 0.5) +
  #Add significance indicators
   geom_point(data = subset(res_df, Significant == "Yes"),
              aes(x = ifelse(log2FoldChange > 0, 
                              log2FoldChange + max(abs(log2FoldChange))*0.05,
                              log2FoldChange - max(abs(log2FoldChange))*0.05)),
             shape = 8, size = 2, color = "black", show.legend = FALSE) +
  theme_minimal(base_size = 12) +
    theme(
      plot.title = element_text(face = "bold", hjust = 0.5, size = 12),
      plot.subtitle = element_text(hjust = 0.5, color = "gray40", size = 12),
      axis.title = element_text(face = "bold"),
      axis.text.y = element_text(size = 11),
      axis.text.x = element_text(size = 11),
      legend.position = "bottom",
      legend.title = element_text(face = "bold"),
      panel.grid.major.y = element_blank(),
      panel.grid.minor.y = element_blank(),
      panel.grid.major.x = element_line(color = "gray90"),
      panel.grid.minor.x = element_blank(),
      plot.caption = element_text(color = "gray50", size = 9)
    )

p5_3
```


***
Let's try to fit Iteration 1 (After 3 months) as a panel using Cowplots:

```{r}
library(cowplot)

p1_2_modified = p1_2 + 
  theme(legend.position = "none")+
  labs(y = NULL)

p1_3_modified <- p1_3 + 
  theme(legend.position = "none")+
  labs(y = NULL)

# Align plots vertically with same width
aligned_plots <- align_plots(p1_1,p1_2_modified,p1_3_modified, align = "v", axis = "l", greedy = F)

# Plot grid with equal widths
plot_i1 = plot_grid(plotlist = aligned_plots, 
          ncol = 3, 
          nrow = 1,
          align = "h",
          rel_heights = c(1, 1, 1))  # Adjust based on number of plots
plot_i1

#ggsave("plot_i1.tiff", plot = plot_i1, width = 16, height = 8, units = "in", dpi = 300)

```


Let's try the same with Iteration 3: 

```{r}

p3_2_modified = p3_2 + 
  theme(legend.position = "none")+
  labs(y = NULL)

p3_3_modified <- p3_3 + 
  theme(legend.position = "none")+
  labs(y = NULL)

# Align plots vertically with same width
aligned_plots <- align_plots(p3_1,p3_2_modified,p3_3_modified, align = "v", axis = "l", greedy = F)

# Plot grid with equal widths
plot_i3 = plot_grid(plotlist = aligned_plots, 
          ncol = 3, 
          nrow = 1,
          align = "h",
          rel_heights = c(1, 1, 1))  # Adjust based on number of plots
plot_i3

#ggsave("plot_i3.tiff", plot = plot_i3, width = 16, height = 8, units = "in", dpi = 300)
```


Let's try the same with Iteration 5: 

```{r}

p5_2_modified = p5_2 + 
  theme(legend.position = "none")+
  labs(y = NULL)

p5_3_modified <- p5_3 + 
  theme(legend.position = "none")+
  labs(y = NULL)

# Align plots vertically with same width
aligned_plots <- align_plots(p5_1,p5_2_modified,p5_3_modified, align = "v", axis = "l", greedy = F)

# Plot grid with equal widths
plot_i5 = plot_grid(plotlist = aligned_plots, 
          ncol = 3, 
          nrow = 1,
          align = "h",
          rel_heights = c(1, 1, 1))  # Adjust based on number of plots
plot_i5

#ggsave("plot_i5.tiff", plot = plot_i5, width = 16, height = 8, units = "in", dpi = 300)
```


```{r}
plot_Fig3 = plot_grid(plotlist = list(plot_i1,plot_i3,plot_i5), 
          nrow = 3,
          align = "h",
          labels = c("A", "B", "C"),
          rel_heights = c(1, 1, 1))  # Adjust based on number of plots
plot_Fig3
```

```{r}
#ggsave("plot_Fig4.tiff", plot = plot_Fig3, width = 14, height = 12, units = "in", dpi = 600)
```


