---
title: "R Notebook"
output: html_notebook
---

```{r}
library(vegan)
library(tidyverse)
library(phyloseq)
library(ggplot2)
library(RColorBrewer)
library(DESeq2)
library(reshape2)
library(emmeans)
library(lme4)
library(lmerTest)
```


```{r}
ps_merged <- readRDS("~/Library/CloudStorage/Box-Box/Lankau_Lab/Greenhouse_Exmpt_2018/16S/Alpha_div/ps_merged.RDS")
physeq_filtered = subset_samples(ps_merged, Soil_inoculation != "Sterile" & Response != "NA")

```

**Function to calculate Alpha Diversity:**
```{r}
alpha_div_f = function(psdata,VALUE){
  psdata = subset_samples(psdata)
  psdata = tax_glom(psdata,taxrank=rank_names(psdata)[VALUE])
  alpha_div = estimate_richness(psdata,measures = c("Chao1","Shannon", "InvSimpson"))
  data_taxa = sample_data(psdata)
  merged_data_taxa = merge(data_taxa,alpha_div, by = "row.names")
  return(merged_data_taxa)
}
```


```{r}

merged_data_Genus = alpha_div_f(physeq_filtered,6)
#merged_data_Genus = subset(merged_data_Genus, Chao1 != 0 & Simpson != 0)

#merged_data_fam = alpha_div_f(physeq_trimmed,5)
#merged_data_ASV = subset(merged_data_ASV, Chao1 != 0 & Simpson != 0)

merged_data_Phyla = alpha_div_f(physeq_filtered,2)
#merged_data_Phyla = subset(merged_data_Phyla, Chao1 != 0& Simpson != 0)

```
*Note*: Ofc there's no singletons because DADA2 removes singletons from our dataset. We deliberately trimmed low abundance because it's probably just noise or is an artifact potentially.   

#Modeling Selection and Diversity:

```{r}
merged_data_Genus1 = subset(merged_data_Genus, Selection.x != "none")
merged_data_Phyla1 = subset(merged_data_Phyla, Selection.x != "none")

#Export Phyla alpha diversity table for PSF analysis GH25_2020
#setwd("~/Library/CloudStorage/Box-Box/Lankau_Lab/Greenhouse_Exmpt_2020")
#write.csv(merged_data_Phyla1, "merged_data_Phyla1.csv", row.names = F)

merged_data_Genus1$Iteration = as.numeric(merged_data_Genus1$Iteration)
merged_data_Phyla1$Iteration = as.numeric(merged_data_Phyla1$Iteration)

model_Chao1 = lmer(Chao1 ~ Clade * Soil_type.x * Selection.x * Iteration + (1|Iteration/Block) + (1|Clade/Genotype), data = subset(merged_data_Phyla1, Selection.x != "none"))
anova(model_Chao1)
summary(model_Chao1)

```
We do see Selection having an effect on Diversity! 


```{r}
model_Shannon = lmer(Shannon ~ Clade * Soil_type.x * Selection.x * Iteration + (1|Iteration/Block) + (1|Clade/Genotype), data = merged_data_Genus1)
anova(model_Shannon)
summary(model_Shannon)
```
Selection seems to have an effect on the variance but not the average in Shannon diversity...


```{r}
library(emmeans)
emmeans_Shannon_Select = emmeans(model_Shannon, "Selection.x", by = "Iteration")
pairs(emmeans_Shannon_Select)
```
Positively selected soils overall over the levels of Clade and Soil_type contain greater alpha diversity. 

ANOVA table tells us that Selection, Iteration, and Clade had an effect on variance in shannon diveristy. We saw a Clade:Iteration, and Selection:Iteration interaction effects.

```{r}

model_Shannon1 = lmer(Shannon ~ Soil_type.x * Selection.x * Iteration + (1|Iteration/Block) + (1|Genotype), data = subset(merged_data_Genus1, Clade == "M"))
AOV_Shannon1 = anova(model_Shannon1)
sum_Shannon1 = summary(model_Shannon1)
Rsq_Shannon1 = r.squaredGLMM(model_Shannon1)

model_Shannon2 = lmer(Shannon ~ Soil_type.x * Selection.x * Iteration + (1|Iteration/Block) + (1|Genotype), data = subset(merged_data_Genus1, Clade == "C"))
AOV_Shannon2= anova(model_Shannon2)
sum_Shannon2 = summary(model_Shannon2)
Rsq_Shannon2 = r.squaredGLMM(model_Shannon2)

model_Shannon3 = lmer(Shannon ~ Soil_type.x * Selection.x * Iteration + (1|Iteration/Block) + (1|Genotype), data = subset(merged_data_Genus1, Clade == "A"))
AOV_Shannon3 = anova(model_Shannon3)
sum_Shannon3 = summary(model_Shannon3)
Rsq_Shannon3 = r.squaredGLMM(model_Shannon3)

```

```{r}
merged_data_Genus$Iteration = as.factor(merged_data_Genus$Iteration)

ggplot_shannon= ggplot(merged_data_Genus, aes(x = Iteration, y = Shannon, fill = Selection.x))+
  geom_boxplot(position=position_dodge(1))+
  geom_dotplot(binaxis = "y", stackdir = "center", alpha = 2, dotsize = 0.5, position = position_dodge(1))+
  theme_bw()+
  facet_wrap(~Clade)

ggplot_shannon
```




```{r}
merged_data_Genus$Iteration = as.factor(merged_data_Genus$Iteration)

merged_data_GenusH = subset(merged_data_Genus, Soil_type.x == "Hancock")

ggplot_shannonH= ggplot(merged_data_GenusH, aes(x = Iteration, y = Shannon, fill = Selection.x))+
  geom_boxplot(position=position_dodge(1))+
  geom_dotplot(binaxis = "y", stackdir = "center", alpha = 2, dotsize = 0.5, position = position_dodge(1))+
  theme_bw()+
  facet_wrap(~Clade)

ggplot_shannonH
```


```{r}
merged_data_Genus$Iteration = as.factor(merged_data_Genus$Iteration)

merged_data_GenusW = subset(merged_data_Genus, Soil_type.x == "Winterfell")

ggplot_shannonW= ggplot(merged_data_GenusW, aes(x = Iteration, y = Shannon, fill = Selection.x))+
  geom_boxplot(position=position_dodge(1))+
  geom_dotplot(binaxis = "y", stackdir = "center", alpha = 2, dotsize = 0.5, position = position_dodge(1))+
  theme_bw()+
  facet_wrap(~Clade)

ggplot_shannonW
```



```{r}
model_InvSimpson1 = lmer(InvSimpson ~ Soil_type.x * Selection.x * Iteration + (1|Iteration/Block) + (1|Genotype), data = subset(merged_data_Phyla1, Clade == "M"))
AOV_InvSimpson1 = anova(model_InvSimpson1)
sum_InvSimpson1 = summary(model_InvSimpson1)
#Rsq_InvSimpson1 = r.squaredGLMM(model_InvSimpson1)

model_InvSimpson2 = lmer(InvSimpson ~ Soil_type.x * Selection.x * Iteration + (1|Iteration/Block) + (1|Genotype), data = subset(merged_data_Phyla1, Clade == "C"))
AOV_InvSimpson2= anova(model_InvSimpson2)
sum_InvSimpson2 = summary(model_InvSimpson2)
#Rsq_InvSimpson2 = r.squaredGLMM(model_InvSimpson2)

model_InvSimpson3 = lmer(InvSimpson ~ Soil_type.x * Selection.x * Iteration + (1|Iteration/Block) + (1|Genotype), data = subset(merged_data_Phyla1, Clade == "A"))
AOV_InvSimpson3 = anova(model_InvSimpson3)
sum_InvSimpson3 = summary(model_InvSimpson3)
#Rsq_InvSimpson3 = r.squaredGLMM(model_InvSimpson3)
```

```{r}
sum_InvSimpson2
```



```{r}
merged_data_Phyla1$Iteration = as.factor(merged_data_Phyla1$Iteration)

merged_data_PhylaW = subset(merged_data_Phyla1, Soil_type.x == "Winterfell")

ggplot_invsimp= ggplot(merged_data_PhylaW, aes(x = Iteration, y = InvSimpson, fill = Selection.x))+
  geom_boxplot(position=position_dodge(1))+
  geom_dotplot(binaxis = "y", stackdir = "center", alpha = 2, dotsize = 0.5, position = position_dodge(1))+
  theme_bw()+
  facet_wrap(~Clade)

ggplot_invsimp
```


```{r}
merged_data_Phyla1$Iteration = as.factor(merged_data_Phyla1$Iteration)

merged_data_PhylaH = subset(merged_data_Phyla1, Soil_type.x == "Hancock")

ggplot_invsimp= ggplot(merged_data_PhylaH, aes(x = Iteration, y = InvSimpson, fill = Selection.x))+
  geom_boxplot(position=position_dodge(1))+
  geom_dotplot(binaxis = "y", stackdir = "center", alpha = 2, dotsize = 0.5, position = position_dodge(1))+
  theme_bw()+
  facet_wrap(~Clade)

ggplot_invsimp
```


#Modeling Genus Alpha Diversity and Response at Iteration 5: 


**SI TABLE for Ch3**
```{r}
InvSimpsonA = lmer(Response ~  InvSimpson * Soil_type.x * Selection.x + (1|Genotype) + (1|Block), data = subset(merged_data_Phyla, Iteration == 5 & Clade == "A"))
sum_InvSimpsonA  = summary(InvSimpsonA)
AOV_InvSimpsonA = anova(InvSimpsonA)

InvSimpsonC = lmer(Response ~  Chao1 * Soil_type.x * Selection.x + (1|Genotype) + (1|Block), data = subset(merged_data_Phyla, Iteration == 5 & Clade == "C"))
sum_InvSimpsonC  = summary(InvSimpsonC)
AOV_InvSimpsonC = anova(InvSimpsonC)

InvSimpsonM = lmer(Response ~  InvSimpson * Soil_type.x * Selection.x + (1|Genotype) + (1|Block) , data = subset(merged_data_Phyla, Iteration == 5 & Clade == "M"))
sum_InvSimpsonM  = summary(InvSimpsonM)
AOV_InvSimpsonM = anova(InvSimpsonM)
```

```{r}
sum_InvSimpsonM
```


```{r}
sum_InvSimpsonC
```



```{r}
InvSimpsonM = lmer(Response ~  Shannon * Soil_type.x * Selection.x + (1|Genotype) + (1|Block) , data = subset(merged_data_Genus, Iteration == 5 & Clade == "M"))
sum_InvSimpsonM  = summary(InvSimpsonM)
AOV_InvSimpsonM = anova(InvSimpsonM)
```

